% Files used
% BurkertRobertsW.tex
% ^^ control ^^
%---
%Length. There is a 50-page limit on the length of regular reviews and a 20-page limit on colloquia.
%% Holt-Roberts RMP => 133 pages (incl. refs) = 54 journal pages => 2.5 input -> 1 page output.
%% References are not included in the 20,000 words.  If one page = 1000 words, then 16,000 words = approximately 40 pages and 20,000 = 50 pages.
%% So, 40-50 pages in the format established below.
%---
%\documentclass[rmp,aps,preprint,nofootinbib,endfloats]{revtex4}
\documentclass[rmp, aps, floatfix, nofootinbib, twocolumn]{revtex4-1}
%\documentclass[rmp, aps, floatfix, nofootinbib, preprint]{revtex4-1}
%\documentclass[rmp,aps,floatfix,nofootinbib]{revtex4}
%\documentclass[rmp,aps,nofootinbib]{revtex4}

\usepackage{amsmath,amsfonts,amssymb,bm}
\usepackage{dsfont}
\usepackage{graphicx}
\usepackage{color}
\usepackage{soul}
\usepackage{slashed}

\renewcommand{\theequation}{\thesection.\arabic{equation}}
\renewcommand{\thefigure}{\thesection.\arabic{figure}}
\renewcommand{\thetable}{\thesection.\arabic{table}}

% Author-defined keyboard shortcuts. All are simple substitutions.
\def\inbar{\,\vrule height1.5ex width.4pt depth0pt}
\def\IR{\relax{\rm I\kern-.18em R}}
\def\IC{\relax\hbox{$\inbar\kern-.3em{\rm C}$}}
\def\Nc{Noncommutative}
\def\nc{noncommutative}
\def\con{conventional} % and not ``commutative''
\def\Tr{\textrm{Tr}~}
\def\zb{\bar{z}}

\newcommand{\sfrac}[2]{\mbox{\footnotesize $\frac{#1}{#2}$}}

\usepackage[mathscr,scaled=1.15]{urwchancal}
\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}%
{<-> s * [1.15] pzcmi7t}{}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\definecolor{purple}{rgb}{0.5,0,0.5}
\definecolor{blue}{rgb}{0.0,0,0.9}
\definecolor{prdblue}{rgb}{0.133,0.118,0.498}
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=prdblue, citecolor= prdblue, urlcolor=prdblue]{hyperref}


\begin{document}
%\title{Distribution Functions of the Nucleon and Pion in the Valence Region}
%\title{Roper Resonance: \\ After fifty years, a solution to the puzzle}
\title{Roper resonance -- solution to the fifty year puzzle}

\author{Volker D. Burkert}
\email{burkert@jlab.org}
\affiliation{Thomas Jefferson National Accelerator Facility, Newport News, Virginia 23606, USA}

\author{Craig D. Roberts}
\email{cdroberts@anl.gov}
\affiliation{Physics Division, Argonne National Laboratory, Argonne, Illinois 60439, USA}

\renewcommand{\baselinestretch}{1.1}

\begin{abstract}
%\begin{footnotesize}
%% EBAC website http://ebac-theory.jlab.org/
%% Based on text sent to Roy Holt 29 June 2010 at 08:40
For half a century, the Roper resonance has defied understanding.  Discovered in 1963, it appears to be an exact copy of the proton \emph{except} that its mass is 50\% greater.
%
The mass is the first problem: it is difficult to explain with any theoretical tool that can validly be used to study the strong-interaction piece of the Standard Model of Particle Physics, \emph{i.e}.\ quantum chromodynamics [QCD].
%
%%https://www.jlab.org/25years/stories/Story6GeV-CEBAF.pdf ... getting JLab to 6 GeV
In the last decade, a new challenge has appeared, \emph{viz}.\ precise information on the proton-to-Roper electroproduction transition form factors, reaching out to momentum transfer $Q^2 \approx 4.5\,$GeV$^2$.  This scale probes the domain within which hard valence-quark degrees-of-freedom could be expected to determine form factor behavior.  Hence, with this new data the Roper resonance becomes a problem for strong-QCD [sQCD].
%
An explanation of how and where the Roper resonance fits into the emerging spectrum of hadrons cannot rest on a description of its mass alone.  Instead, it must combine an understanding of the Roper's mass and width with a detailed account of its structure and how that structure is revealed in the momentum dependence of the transition form factors.  Furthermore, it must unify all this with a similarly complete picture of the proton from which the Roper resonance is produced.
%
This is a prodigious task, but a ten-year international collaborative effort, drawing together experimentalists and theorists, has presented a solution to the puzzle.  Namely, the observed Roper is at heart the proton's first radial excitation, consisting of a dressed-quark core augmented by a meson cloud that reduces the core mass by approximately 20\% and materially alters its electroproduction form factors on $Q^2 < 2\,m_N^2$, where $m_N$ is the proton's mass.
%
We describe the experimental motivations and developments which enabled electroproduction data to be procured within a domain that is unambiguously the purview of sQCD, thereby providing a real challenge and opportunity for modern theory; and survey the developments in reaction models and QCD theory that have enabled this conclusion to be drawn about the nature of the Roper resonance.
%
%\end{footnotesize}
\end{abstract}
\date{21 September 2017}
%\date{01 August 2017}
%\date{29 June 2017}
%\date{8 August 2016}

\maketitle

%\renewcommand{\baselinestretch}{1.6}

\tableofcontents
%\newpage

\input 1Intro
%\newpage

\input 2CQM
%\newpage

\input 3Experiment
%\newpage

\input 4DCC
%\newpage

\input 5Theory
%\newpage

\input 6Summary

%

\section*{Acknowledgments}
In preparing this article we benefited greatly from constructive comments and input provided by
%
I.\,G.~Aznauryan,
%
A.~Bashir,
%
D.~Binosi,
%
L.~Chang,
%
C.~Chen,
%
Z.-F.~Cui,
%
B.~El-Bennich,
%
R.~Gothe,
%
L.\,X.~Guti\'errez-Guerrero,
%
G.~Krein,
%
T.-S.\,H.~Lee,
%
H.-W.~Lin,
%
K.-F.~Liu,
%
Y.~Lu,
%
C.~Mezrag,
%
V.~Mokeev,
%
J.~Papavassiliou,
%
J.-L.~Ping,
%
H.\,L.\,L.~Roberts,
%
J.~Rodrig\'{\i}uez-Quintero,
%
T.~Sato,
%
S.\,M.~Schmidt,
%
J.~Segovia,
%
S.-S.~Xu,
%
F.~Wang,
%
D.\,J.~Wilson,
%
and H.-S.~Zong.
%
This work was supported by
%
U.S.\ Department of Energy, Office of Science, Office of Nuclear Physics, under contract nos.~DE-AC05-06OR23177 and DE-AC02-06CH11357.

%\newpage

%\input 1607.04536

%\newpage

%\input references
%%\bibliographystyle{apsrmp}
%\bibliographystyle{../../../../zProc/z10/z10KITPC/h-physrev4}
%... 223 references ... and counting.
%%\bibliography{../../../CollectedBiB}

\begin{thebibliography}{227}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname bibnamefont\endcsname\relax
  \def\bibnamefont#1{#1}\fi
\expandafter\ifx\csname bibfnamefont\endcsname\relax
  \def\bibfnamefont#1{#1}\fi
\expandafter\ifx\csname citenamefont\endcsname\relax
  \def\citenamefont#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{Aaij \emph{et~al.}(2015)\citenamefont{Aaij}
  \emph{et~al.}}]{Aaij:2015tga}
\bibinfo{author}{\bibnamefont{Aaij}, \bibfnamefont{R.}}, \emph{et~al.},
  \bibinfo{year}{2015}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{072001}.

\bibitem[{\citenamefont{Abbott}(1981)}]{Abbott:1980hw}
\bibinfo{author}{\bibnamefont{Abbott}, \bibfnamefont{L.~F.}},
  \bibinfo{year}{1981}, \bibinfo{journal}{Nucl. Phys. B}
  \textbf{\bibinfo{volume}{185}}, \bibinfo{pages}{189}.

\bibitem[{\citenamefont{Abbott}(1982)}]{Abbott:1981ke}
\bibinfo{author}{\bibnamefont{Abbott}, \bibfnamefont{L.~F.}},
  \bibinfo{year}{1982}, \bibinfo{journal}{Acta Phys. Polon. B}
  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{33}.

\bibitem[{Abe \emph{et~al.}(1995{\natexlab{a}})\citenamefont{Abe}
  \emph{et~al.}}]{Abe:1995rn}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1995}{\natexlab{a}}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{364}}, \bibinfo{pages}{61}.

\bibitem[{Abe \emph{et~al.}(1995{\natexlab{b}})\citenamefont{Abe}
  \emph{et~al.}}]{Abe:1995mt}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1995}{\natexlab{b}}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{75}}, \bibinfo{pages}{25}.

\bibitem[{Abe \emph{et~al.}(1995{\natexlab{c}})\citenamefont{Abe}
  \emph{et~al.}}]{Abe:1994cp}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1995}{\natexlab{c}}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{74}}, \bibinfo{pages}{346}.

\bibitem[{Abe \emph{et~al.}(1996)\citenamefont{Abe} \emph{et~al.}}]{Abe:1995dc}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1996}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{76}}, \bibinfo{pages}{587}.

\bibitem[{Abe \emph{et~al.}(1997{\natexlab{a}})\citenamefont{Abe}
  \emph{et~al.}}]{Abe:1997qk}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1997}{\natexlab{a}}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{404}}, \bibinfo{pages}{377}.

\bibitem[{Abe \emph{et~al.}(1997{\natexlab{b}})\citenamefont{Abe}
  \emph{et~al.}}]{Abe:1997dp}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1997}{\natexlab{b}}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{405}}, \bibinfo{pages}{180}.

\bibitem[{Abe \emph{et~al.}(1997{\natexlab{c}})\citenamefont{Abe}
  \emph{et~al.}}]{Abe:1997cx}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1997}{\natexlab{c}}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{79}}, \bibinfo{pages}{26}.

\bibitem[{Abe \emph{et~al.}(1998)\citenamefont{Abe} \emph{et~al.}}]{Abe:1998wq}
\bibinfo{author}{\bibnamefont{Abe}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1998}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{D58}}, \bibinfo{pages}{112003}.

\bibitem[{Ackerstaff \emph{et~al.}(1997)\citenamefont{Ackerstaff}
  \emph{et~al.}}]{Ackerstaff:1997ws}
\bibinfo{author}{\bibnamefont{Ackerstaff}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1997}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{404}}, \bibinfo{pages}{383}.

\bibitem[{Ackerstaff \emph{et~al.}(1998)\citenamefont{Ackerstaff}
  \emph{et~al.}}]{Ackerstaff:1998ja}
\bibinfo{author}{\bibnamefont{Ackerstaff}, \bibfnamefont{K.}}, \emph{et~al.},
  \bibinfo{year}{1998}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{444}}, \bibinfo{pages}{531}.

\bibitem[{\citenamefont{Adelman}(1964)}]{PhysRevLett.13.555}
\bibinfo{author}{\bibnamefont{Adelman}, \bibfnamefont{S.~L.}},
  \bibinfo{year}{1964}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{555}.

\bibitem[{Adolph \emph{et~al.}(2016)\citenamefont{Adolph}
  \emph{et~al.}}]{Adolph:2015saz}
\bibinfo{author}{\bibnamefont{Adolph}, \bibfnamefont{C.}}, \emph{et~al.},
  \bibinfo{year}{2016}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{753}}, \bibinfo{pages}{18}.

\bibitem[{\citenamefont{Aguilar} \emph{et~al.}(2016)\citenamefont{Aguilar,
  Binosi, and Papavassiliou}}]{Aguilar:2015bud}
\bibinfo{author}{\bibnamefont{Aguilar}, \bibfnamefont{A.~C.}},
  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Binosi}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}},
  \bibinfo{year}{2016}, \bibinfo{journal}{Front. Phys. China}
  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{111203}.

\bibitem[{Airapetian \emph{et~al.}(1998)\citenamefont{Airapetian}
  \emph{et~al.}}]{Airapetian:1998wi}
\bibinfo{author}{\bibnamefont{Airapetian}, \bibfnamefont{A.}}, \emph{et~al.},
  \bibinfo{year}{1998}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{442}}, \bibinfo{pages}{484}.

\bibitem[{Airapetian \emph{et~al.}(2003)\citenamefont{Airapetian}
  \emph{et~al.}}]{Airapetian:2002rw}
\bibinfo{author}{\bibnamefont{Airapetian}, \bibfnamefont{A.}}, \emph{et~al.},
  \bibinfo{year}{2003}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{90}}, \bibinfo{pages}{092002}.

\bibitem[{Airapetian \emph{et~al.}(2007)\citenamefont{Airapetian}
  \emph{et~al.}}]{Airapetian:2006vy}
\bibinfo{author}{\bibnamefont{Airapetian}, \bibfnamefont{A.}}, \emph{et~al.},
  \bibinfo{year}{2007}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{75}}, \bibinfo{pages}{012007}.

\bibitem[{Alekseev \emph{et~al.}(2010)\citenamefont{Alekseev}
  \emph{et~al.}}]{Alekseev:2010hc}
\bibinfo{author}{\bibnamefont{Alekseev}, \bibfnamefont{M.~G.}}, \emph{et~al.},
  \bibinfo{year}{2010}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{690}}, \bibinfo{pages}{466}.

\bibitem[{Alexakhin \emph{et~al.}(2007)\citenamefont{Alexakhin}
  \emph{et~al.}}]{Alexakhin:2006oza}
\bibinfo{author}{\bibnamefont{Alexakhin}, \bibfnamefont{V.~{\relax Yu}.}},
  \emph{et~al.}, \bibinfo{year}{2007}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{647}}, \bibinfo{pages}{8}.

\bibitem[{\citenamefont{Alexandrou}
  \emph{et~al.}(2006)\citenamefont{Alexandrou, de~Forcrand, and
  Lucini}}]{Alexandrou:2006cq}
\bibinfo{author}{\bibnamefont{Alexandrou}, \bibfnamefont{C.}},
  \bibinfo{author}{\bibfnamefont{{\mbox{Ph}}.}~\bibnamefont{de~Forcrand}}, and
  \bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{Lucini}},
  \bibinfo{year}{2006}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{97}}, \bibinfo{pages}{222002}.

\bibitem[{\citenamefont{Alexandrou}
  \emph{et~al.}(2015)\citenamefont{Alexandrou, Leontiou, Papanicolas, and
  Stiliaris}}]{Alexandrou:2014mka}
\bibinfo{author}{\bibnamefont{Alexandrou}, \bibfnamefont{C.}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Leontiou}},
  \bibinfo{author}{\bibfnamefont{C.~N.} \bibnamefont{Papanicolas}}, and
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Stiliaris}},
  \bibinfo{year}{2015}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{91}}, \bibinfo{pages}{014506}.

\bibitem[{\citenamefont{Alexandrou}
  \emph{et~al.}(2012)\citenamefont{Alexandrou, Papanicolas, and
  Vanderhaeghen}}]{Alexandrou:2012da}
\bibinfo{author}{\bibnamefont{Alexandrou}, \bibfnamefont{C.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Papanicolas}}, and
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Vanderhaeghen}},
  \bibinfo{year}{2012}, \bibinfo{journal}{Rev. Mod. Phys.}
  \textbf{\bibinfo{volume}{84}}, \bibinfo{pages}{1231}.

\bibitem[{\citenamefont{Altarelli and Parisi}(1977)}]{Altarelli:1977}
\bibinfo{author}{\bibnamefont{Altarelli}, \bibfnamefont{G.}}, and
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Parisi}},
  \bibinfo{year}{1977}, \bibinfo{journal}{Nucl. Phys. B}
  \textbf{\bibinfo{volume}{126}}, \bibinfo{pages}{298}.

\bibitem[{Anisovich \emph{et~al.}(2017)\citenamefont{Anisovich}
  \emph{et~al.}}]{Anisovich:2017pmi}
\bibinfo{author}{\bibnamefont{Anisovich}, \bibfnamefont{A.~V.}}, \emph{et~al.},
  \bibinfo{year}{2017}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{119}}, \bibinfo{pages}{062004}.

\bibitem[{Anthony \emph{et~al.}(1993)\citenamefont{Anthony}
  \emph{et~al.}}]{Anthony:1993uf}
\bibinfo{author}{\bibnamefont{Anthony}, \bibfnamefont{P.~L.}}, \emph{et~al.},
  \bibinfo{year}{1993}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{71}}, \bibinfo{pages}{959}.

\bibitem[{Anthony \emph{et~al.}(1996)\citenamefont{Anthony}
  \emph{et~al.}}]{Anthony:1996mw}
\bibinfo{author}{\bibnamefont{Anthony}, \bibfnamefont{P.~L.}}, \emph{et~al.},
  \bibinfo{year}{1996}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{54}}, \bibinfo{pages}{6620}.

\bibitem[{Anthony \emph{et~al.}(1999{\natexlab{a}})\citenamefont{Anthony}
  \emph{et~al.}}]{Anthony:1999rm}
\bibinfo{author}{\bibnamefont{Anthony}, \bibfnamefont{P.~L.}}, \emph{et~al.},
  \bibinfo{year}{1999}{\natexlab{a}}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{463}}, \bibinfo{pages}{339}.

\bibitem[{Anthony \emph{et~al.}(1999{\natexlab{b}})\citenamefont{Anthony}
  \emph{et~al.}}]{Anthony:1999py}
\bibinfo{author}{\bibnamefont{Anthony}, \bibfnamefont{P.~L.}}, \emph{et~al.},
  \bibinfo{year}{1999}{\natexlab{b}}, \bibinfo{journal}{Phys. Lett.}
  \textbf{\bibinfo{volume}{B458}}, \bibinfo{pages}{529}.

\bibitem[{Anthony \emph{et~al.}(2000)\citenamefont{Anthony}
  \emph{et~al.}}]{Anthony:2000fn}
\bibinfo{author}{\bibnamefont{Anthony}, \bibfnamefont{P.~L.}}, \emph{et~al.},
  \bibinfo{year}{2000}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{493}}, \bibinfo{pages}{19}.

\bibitem[{Anthony \emph{et~al.}(2003)\citenamefont{Anthony}
  \emph{et~al.}}]{Anthony:2002hy}
\bibinfo{author}{\bibnamefont{Anthony}, \bibfnamefont{P.~L.}}, \emph{et~al.},
  \bibinfo{year}{2003}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{553}}, \bibinfo{pages}{18}.

\bibitem[{\citenamefont{Aoki} \emph{et~al.}(2013)\citenamefont{Aoki, Aoyama,
  Kurachi, Maskawa, Nagai, Ohki, Shibata, Yamawaki, and
  Yamazaki}}]{Aoki:2013xza}
\bibinfo{author}{\bibnamefont{Aoki}, \bibfnamefont{Y.}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Aoyama}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Kurachi}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Maskawa}},
  \bibinfo{author}{\bibfnamefont{K.-i.} \bibnamefont{Nagai}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Ohki}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Shibata}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Yamawaki}}, and
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Yamazaki}},
  \bibinfo{year}{2013}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{87}}, \bibinfo{pages}{094511}.

\bibitem[{\citenamefont{Appelquist}
  \emph{et~al.}(1996)\citenamefont{Appelquist, Terning, and
  Wijewardhana}}]{Appelquist:1996dq}
\bibinfo{author}{\bibnamefont{Appelquist}, \bibfnamefont{T.}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Terning}}, and
  \bibinfo{author}{\bibfnamefont{L.~C.~R.} \bibnamefont{Wijewardhana}},
  \bibinfo{year}{1996}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{1214}.

\bibitem[{Appelquist \emph{et~al.}(2010)\citenamefont{Appelquist}
  \emph{et~al.}}]{Appelquist:2009ka}
\bibinfo{author}{\bibnamefont{Appelquist}, \bibfnamefont{T.}}, \emph{et~al.},
  \bibinfo{year}{2010}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{104}}, \bibinfo{pages}{071601}.

\bibitem[{\citenamefont{Arndt} \emph{et~al.}(2006)\citenamefont{Arndt, Briscoe,
  Strakovsky, and Workman}}]{Arndt:2006bf}
\bibinfo{author}{\bibnamefont{Arndt}, \bibfnamefont{R.~A.}},
  \bibinfo{author}{\bibfnamefont{W.~J.} \bibnamefont{Briscoe}},
  \bibinfo{author}{\bibfnamefont{I.~I.} \bibnamefont{Strakovsky}}, and
  \bibinfo{author}{\bibfnamefont{R.~L.} \bibnamefont{Workman}},
  \bibinfo{year}{2006}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{74}}, \bibinfo{pages}{045205}.

\bibitem[{\citenamefont{Arndt} \emph{et~al.}(1985)\citenamefont{Arndt, Ford,
  and Roper}}]{Arndt:1985vj}
\bibinfo{author}{\bibnamefont{Arndt}, \bibfnamefont{R.~A.}},
  \bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Ford}}, and
  \bibinfo{author}{\bibfnamefont{L.~D.} \bibnamefont{Roper}},
  \bibinfo{year}{1985}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{32}}, \bibinfo{pages}{1085}.

\bibitem[{\citenamefont{Auvil} \emph{et~al.}(1964)\citenamefont{Auvil,
  Lovelace, Donnachie, and Lea}}]{AUVIL196476}
\bibinfo{author}{\bibnamefont{Auvil}, \bibfnamefont{P.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Lovelace}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Donnachie}}, and
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Lea}}, \bibinfo{year}{1964},
  \bibinfo{journal}{Physics Letters} \textbf{\bibinfo{volume}{12}},
  \bibinfo{pages}{76}.

\bibitem[{\citenamefont{Aznauryan and
  Burkert}(2012{\natexlab{a}})}]{Aznauryan:2011qj}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.}}, and
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Burkert}},
  \bibinfo{year}{2012}{\natexlab{a}}, \bibinfo{journal}{Prog. Part. Nucl.
  Phys.} \textbf{\bibinfo{volume}{67}}, \bibinfo{pages}{1}.

\bibitem[{\citenamefont{Aznauryan and
  Burkert}(2012{\natexlab{b}})}]{Aznauryan:2012ec}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.}}, and
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Burkert}},
  \bibinfo{year}{2012}{\natexlab{b}}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{85}}, \bibinfo{pages}{055202}.

\bibitem[{Aznauryan \emph{et~al.}(2009)\citenamefont{Aznauryan}
  \emph{et~al.}}]{Aznauryan:2009mx}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.}}, \emph{et~al.},
  \bibinfo{year}{2009}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{80}}, \bibinfo{pages}{055203}.

\bibitem[{Aznauryan \emph{et~al.}(2013)\citenamefont{Aznauryan}
  \emph{et~al.}}]{Aznauryan:2012baS}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.}}, \emph{et~al.},
  \bibinfo{year}{2013}, \bibinfo{journal}{Int. J. Mod. Phys. E}
  \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{1330015}.

\bibitem[{\citenamefont{Aznauryan}(2003)}]{Aznauryan:2002gd}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.~G.}},
  \bibinfo{year}{2003}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{67}}, \bibinfo{pages}{015209}.

\bibitem[{\citenamefont{Aznauryan and Burkert}(2016)}]{Aznauryan:2016wwm}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.~G.}}, and
  \bibinfo{author}{\bibfnamefont{V.~D.} \bibnamefont{Burkert}},
  \bibinfo{year}{2016}, \bibinfo{note}{{\emph{Configuration mixings and
  light-front relativistic quark model predictions for the electroexcitation of
  the \mbox{$\Delta(1232)3/2^+$}, \mbox{$N(1440)1/2^+$}, and
  \mbox{$\Delta(1600)3/2^+$}}, arXiv:1603.06692 [hep-ph]}}.

\bibitem[{\citenamefont{Aznauryan} \emph{et~al.}(2005)\citenamefont{Aznauryan,
  Burkert, Egiyan, Joo, Minehart, and Smith}}]{Aznauryan:2004jd}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.~G.}},
  \bibinfo{author}{\bibfnamefont{V.~D.} \bibnamefont{Burkert}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Egiyan}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Joo}},
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Minehart}}, and
  \bibinfo{author}{\bibfnamefont{L.~C.} \bibnamefont{Smith}},
  \bibinfo{year}{2005}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{71}}, \bibinfo{pages}{015201}.

\bibitem[{Aznauryan \emph{et~al.}(2008)\citenamefont{Aznauryan}
  \emph{et~al.}}]{Aznauryan:2008pe}
\bibinfo{author}{\bibnamefont{Aznauryan}, \bibfnamefont{I.~G.}}, \emph{et~al.},
  \bibinfo{year}{2008}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{78}}, \bibinfo{pages}{045209}.

\bibitem[{\citenamefont{Babich} \emph{et~al.}(2007)\citenamefont{Babich,
  Garron, Hoelbling, Howard, Lellouch} \emph{et~al.}}]{Babich:2007ah}
\bibinfo{author}{\bibnamefont{Babich}, \bibfnamefont{R.}},
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Garron}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Hoelbling}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Howard}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Lellouch}}, \emph{et~al.},
  \bibinfo{year}{2007}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{76}}, \bibinfo{pages}{074021}.

\bibitem[{\citenamefont{Ball and Chiu}(1980)}]{Ball:1980ay}
\bibinfo{author}{\bibnamefont{Ball}, \bibfnamefont{J.~S.}}, and
  \bibinfo{author}{\bibfnamefont{T.-W.} \bibnamefont{Chiu}},
  \bibinfo{year}{1980}, \bibinfo{journal}{Phys.\ Rev.\ D}
  \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{2542}.

\bibitem[{\citenamefont{Bareyre} \emph{et~al.}(1964)\citenamefont{Bareyre,
  Bricman, Valladas, Villet, Bizard, and Seguinot}}]{BAREYRE1964137}
\bibinfo{author}{\bibnamefont{Bareyre}, \bibfnamefont{P.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Bricman}},
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Valladas}},
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Villet}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Bizard}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Seguinot}},
  \bibinfo{year}{1964}, \bibinfo{journal}{Physics Letters}
  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{137}.

\bibitem[{\citenamefont{Barnes and Close}(1983)}]{Barnes:1982fj}
\bibinfo{author}{\bibnamefont{Barnes}, \bibfnamefont{T.}}, and
  \bibinfo{author}{\bibfnamefont{F.~E.} \bibnamefont{Close}},
  \bibinfo{year}{1983}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{123}}, \bibinfo{pages}{89}.

\bibitem[{Bashir \emph{et~al.}(2012)\citenamefont{Bashir}
  \emph{et~al.}}]{Bashir:2012fs}
\bibinfo{author}{\bibnamefont{Bashir}, \bibfnamefont{A.}}, \emph{et~al.},
  \bibinfo{year}{2012}, \bibinfo{journal}{Commun. Theor. Phys.}
  \textbf{\bibinfo{volume}{58}}, \bibinfo{pages}{79}.

\bibitem[{\citenamefont{Bender} \emph{et~al.}(2002)\citenamefont{Bender,
  Detmold, Roberts, and Thomas}}]{Bender:2002as}
\bibinfo{author}{\bibnamefont{Bender}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Detmold}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{A.~W.} \bibnamefont{Thomas}},
  \bibinfo{year}{2002}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{65}}, \bibinfo{pages}{065203}.

\bibitem[{\citenamefont{Bender} \emph{et~al.}(1996)\citenamefont{Bender,
  Roberts, and von Smekal}}]{Bender:1996bb}
\bibinfo{author}{\bibnamefont{Bender}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{von Smekal}},
  \bibinfo{year}{1996}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{380}}, \bibinfo{pages}{7}.

\bibitem[{\citenamefont{Bhagwat} \emph{et~al.}(2004)\citenamefont{Bhagwat,
  H{\"o}ll, Krassnigg, Roberts, and Tandy}}]{Bhagwat:2004hn}
\bibinfo{author}{\bibnamefont{Bhagwat}, \bibfnamefont{M.~S.}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{H{\"o}ll}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Krassnigg}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{2004}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{70}}, \bibinfo{pages}{035205}.

\bibitem[{\citenamefont{Bhagwat} \emph{et~al.}(2003)\citenamefont{Bhagwat,
  Pichowsky, Roberts, and Tandy}}]{Bhagwat:2003vw}
\bibinfo{author}{\bibnamefont{Bhagwat}, \bibfnamefont{M.~S.}},
  \bibinfo{author}{\bibfnamefont{M.~A.} \bibnamefont{Pichowsky}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{2003}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{68}}, \bibinfo{pages}{015203}.

\bibitem[{\citenamefont{Bhagwat and Tandy}(2006)}]{Bhagwat:2006tu}
\bibinfo{author}{\bibnamefont{Bhagwat}, \bibfnamefont{M.~S.}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{2006}, \bibinfo{journal}{AIP Conf. Proc.}
  \textbf{\bibinfo{volume}{842}}, \bibinfo{pages}{225}.

\bibitem[{\citenamefont{Bijker and Santopinto}(2009)}]{Bijker:2009up}
\bibinfo{author}{\bibnamefont{Bijker}, \bibfnamefont{R.}}, and
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Santopinto}},
  \bibinfo{year}{2009}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{80}}, \bibinfo{pages}{065210}.

\bibitem[{\citenamefont{Binosi}
  \emph{et~al.}(2017{\natexlab{a}})\citenamefont{Binosi, Chang, Papavassiliou,
  Qin, and Roberts}}]{Binosi:2016wcx}
\bibinfo{author}{\bibnamefont{Binosi}, \bibfnamefont{D.}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Chang}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}},
  \bibinfo{author}{\bibfnamefont{S.-X.} \bibnamefont{Qin}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2017}{\natexlab{a}}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{031501(R)}.

\bibitem[{\citenamefont{Binosi} \emph{et~al.}(2015)\citenamefont{Binosi, Chang,
  Papavassiliou, and Roberts}}]{Binosi:2014aea}
\bibinfo{author}{\bibnamefont{Binosi}, \bibfnamefont{D.}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Chang}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2015}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{742}}, \bibinfo{pages}{183}.

\bibitem[{\citenamefont{Binosi}
  \emph{et~al.}(2017{\natexlab{b}})\citenamefont{Binosi, Mezrag, Papavassiliou,
  Roberts, and Rodr{\'i}guez-Quintero}}]{Binosi:2016nmeRMP}
\bibinfo{author}{\bibnamefont{Binosi}, \bibfnamefont{D.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Mezrag}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Rodr{\'i}guez-Quintero}},
  \bibinfo{year}{2017}{\natexlab{b}}, \bibinfo{note}{{\emph{Process-independent
  strong running coupling}, arXiv:1612.04835 [nucl-th]}}.

\bibitem[{\citenamefont{Binosi and Papavassiliou}(2002)}]{Binosi:2002ft}
\bibinfo{author}{\bibnamefont{Binosi}, \bibfnamefont{D.}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}},
  \bibinfo{year}{2002}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{66}}, \bibinfo{pages}{111901}.

\bibitem[{\citenamefont{Binosi and Papavassiliou}(2004)}]{Binosi:2003rr}
\bibinfo{author}{\bibnamefont{Binosi}, \bibfnamefont{D.}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}},
  \bibinfo{year}{2004}, \bibinfo{journal}{J. Phys. G}
  \textbf{\bibinfo{volume}{30}}, \bibinfo{pages}{203}.

\bibitem[{\citenamefont{Binosi and Papavassiliou}(2009)}]{Binosi:2009qm}
\bibinfo{author}{\bibnamefont{Binosi}, \bibfnamefont{D.}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}},
  \bibinfo{year}{2009}, \bibinfo{journal}{Phys. Rept.}
  \textbf{\bibinfo{volume}{479}}, \bibinfo{pages}{1}.

\bibitem[{\citenamefont{Binosi}
  \emph{et~al.}(2017{\natexlab{c}})\citenamefont{Binosi, Roberts, and
  Rodriguez-Quintero}}]{Binosi:2016xxu}
\bibinfo{author}{\bibnamefont{Binosi}, \bibfnamefont{D.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Rodriguez-Quintero}},
  \bibinfo{year}{2017}{\natexlab{c}}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{114009}.

\bibitem[{\citenamefont{Bloch}(2002)}]{Bloch:2002eq}
\bibinfo{author}{\bibnamefont{Bloch}, \bibfnamefont{J.~C.~R.}},
  \bibinfo{year}{2002}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{D66}}, \bibinfo{pages}{034032}.

\bibitem[{\citenamefont{Boucaud} \emph{et~al.}(2012)\citenamefont{Boucaud,
  Leroy, Le-Yaouanc, Micheli, Pene, and
  Rodr{\'i}guez-Quintero}}]{Boucaud:2011ug}
\bibinfo{author}{\bibnamefont{Boucaud}, \bibfnamefont{P.}},
  \bibinfo{author}{\bibfnamefont{J.~P.} \bibnamefont{Leroy}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Le-Yaouanc}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Micheli}},
  \bibinfo{author}{\bibfnamefont{O.}~\bibnamefont{Pene}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Rodr{\'i}guez-Quintero}},
  \bibinfo{year}{2012}, \bibinfo{journal}{Few Body Syst.}
  \textbf{\bibinfo{volume}{53}}, \bibinfo{pages}{387}.

\bibitem[{Bowman \emph{et~al.}(2005)\citenamefont{Bowman}
  \emph{et~al.}}]{Bowman:2005vx}
\bibinfo{author}{\bibnamefont{Bowman}, \bibfnamefont{P.~O.}}, \emph{et~al.},
  \bibinfo{year}{2005}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{71}}, \bibinfo{pages}{054507}.

\bibitem[{\citenamefont{{Braaten, E.}}(2016)}]{BraatenFB21}
\bibinfo{author}{\bibnamefont{{Braaten, E.}}}, \bibinfo{year}{2016},
  \bibinfo{journal}{EPJ Web of Conferences} \textbf{\bibinfo{volume}{113}},
  \bibinfo{pages}{01015}.

\bibitem[{\citenamefont{Briceno} \emph{et~al.}(2017)\citenamefont{Briceno,
  Dudek, and Young}}]{Briceno:2017max}
\bibinfo{author}{\bibnamefont{Briceno}, \bibfnamefont{R.~A.}},
  \bibinfo{author}{\bibfnamefont{J.~J.} \bibnamefont{Dudek}}, and
  \bibinfo{author}{\bibfnamefont{R.~D.} \bibnamefont{Young}},
  \bibinfo{year}{2017}, \bibinfo{note}{{\emph{Scattering processes and
  resonances from lattice QCD}, arXiv: 1706.06223 [hep-lat]}}.

\bibitem[{\citenamefont{Brodsky} \emph{et~al.}(2015)\citenamefont{Brodsky,
  Deshpande, Gao, McKeown, Meyer, Meziani, Milner, Qiu, Richards, and
  Roberts}}]{Brodsky:2015aiaRMP}
\bibinfo{author}{\bibnamefont{Brodsky}, \bibfnamefont{S.~J.}},
  \bibinfo{author}{\bibfnamefont{A.~L.} \bibnamefont{Deshpande}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Gao}},
  \bibinfo{author}{\bibfnamefont{R.~D.} \bibnamefont{McKeown}},
  \bibinfo{author}{\bibfnamefont{C.~A.} \bibnamefont{Meyer}},
  \bibinfo{author}{\bibfnamefont{Z.-E.} \bibnamefont{Meziani}},
  \bibinfo{author}{\bibfnamefont{R.~G.} \bibnamefont{Milner}},
  \bibinfo{author}{\bibfnamefont{J.-W.} \bibnamefont{Qiu}},
  \bibinfo{author}{\bibfnamefont{D.~G.} \bibnamefont{Richards}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2015}, \bibinfo{note}{{\emph{QCD and Hadron Physics},
  aXiv:1502.05728 [hep-ph]}}.

\bibitem[{\citenamefont{Burden} \emph{et~al.}(1989)\citenamefont{Burden,
  Cahill, and Praschifka}}]{Burden:1988dt}
\bibinfo{author}{\bibnamefont{Burden}, \bibfnamefont{C.~J.}},
  \bibinfo{author}{\bibfnamefont{R.~T.} \bibnamefont{Cahill}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Praschifka}},
  \bibinfo{year}{1989}, \bibinfo{journal}{Austral. J. Phys.}
  \textbf{\bibinfo{volume}{42}}, \bibinfo{pages}{147}.

\bibitem[{\citenamefont{Burkert}(2012)}]{Burkert:2012ee}
\bibinfo{author}{\bibnamefont{Burkert}, \bibfnamefont{V.~D.}},
  \bibinfo{year}{2012}, \bibinfo{journal}{EPJ Web Conf.}
  \textbf{\bibinfo{volume}{37}}, \bibinfo{pages}{01017}.

\bibitem[{\citenamefont{Burkert and Lee}(2004)}]{Burkert:2004sk}
\bibinfo{author}{\bibnamefont{Burkert}, \bibfnamefont{V.~D.}}, and
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{year}{2004}, \bibinfo{journal}{Int. J. Mod. Phys. E}
  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{1035}.

\bibitem[{\citenamefont{Cahill}(1989)}]{Cahill:1988zi}
\bibinfo{author}{\bibnamefont{Cahill}, \bibfnamefont{R.~T.}},
  \bibinfo{year}{1989}, \bibinfo{journal}{Austral. J. Phys.}
  \textbf{\bibinfo{volume}{42}}, \bibinfo{pages}{171}.

\bibitem[{\citenamefont{Cahill} \emph{et~al.}(1987)\citenamefont{Cahill,
  Roberts, and Praschifka}}]{Cahill:1987qr}
\bibinfo{author}{\bibnamefont{Cahill}, \bibfnamefont{R.~T.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Praschifka}},
  \bibinfo{year}{1987}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{36}}, \bibinfo{pages}{2804}.

\bibitem[{\citenamefont{Cahill} \emph{et~al.}(1989)\citenamefont{Cahill,
  Roberts, and Praschifka}}]{Cahill:1988dx}
\bibinfo{author}{\bibnamefont{Cahill}, \bibfnamefont{R.~T.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Praschifka}},
  \bibinfo{year}{1989}, \bibinfo{journal}{Austral. J. Phys.}
  \textbf{\bibinfo{volume}{42}}, \bibinfo{pages}{129}.

\bibitem[{\citenamefont{Cano and Gonzalez}(1998)}]{Cano:1998wz}
\bibinfo{author}{\bibnamefont{Cano}, \bibfnamefont{F.}}, and
  \bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Gonzalez}},
  \bibinfo{year}{1998}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{431}}, \bibinfo{pages}{270}.

\bibitem[{\citenamefont{Capstick and Isgur}(1986)}]{Capstick:1986bm}
\bibinfo{author}{\bibnamefont{Capstick}, \bibfnamefont{S.}}, and
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Isgur}},
  \bibinfo{year}{1986}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{2809}.

\bibitem[{\citenamefont{Capstick and Keister}(1995)}]{Capstick:1994ne}
\bibinfo{author}{\bibnamefont{Capstick}, \bibfnamefont{S.}}, and
  \bibinfo{author}{\bibfnamefont{B.~D.} \bibnamefont{Keister}},
  \bibinfo{year}{1995}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{51}}, \bibinfo{pages}{3598}.

\bibitem[{\citenamefont{Capstick and Page}(2002)}]{Capstick:2002wm}
\bibinfo{author}{\bibnamefont{Capstick}, \bibfnamefont{S.}}, and
  \bibinfo{author}{\bibfnamefont{P.~R.} \bibnamefont{Page}},
  \bibinfo{year}{2002}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{66}}, \bibinfo{pages}{065204}.

\bibitem[{\citenamefont{Capstick and Roberts}(2000)}]{Capstick:2000qj}
\bibinfo{author}{\bibnamefont{Capstick}, \bibfnamefont{S.}}, and
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Roberts}},
  \bibinfo{year}{2000}, \bibinfo{journal}{Prog. Part. Nucl. Phys.}
  \textbf{\bibinfo{volume}{45}}, \bibinfo{pages}{S241}.

\bibitem[{\citenamefont{Cardarelli}
  \emph{et~al.}(1997)\citenamefont{Cardarelli, Pace, Salme, and
  Simula}}]{Cardarelli:1996vn}
\bibinfo{author}{\bibnamefont{Cardarelli}, \bibfnamefont{F.}},
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Pace}},
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Salme}}, and
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Simula}},
  \bibinfo{year}{1997}, \bibinfo{journal}{Phys. Lett.}
  \textbf{\bibinfo{volume}{B397}}, \bibinfo{pages}{13}.

\bibitem[{\citenamefont{Carlson}(1986)}]{Carlson:1985mm}
\bibinfo{author}{\bibnamefont{Carlson}, \bibfnamefont{C.~E.}},
  \bibinfo{year}{1986}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{2704}.

\bibitem[{Carman \emph{et~al.}(2014)\citenamefont{Carman}
  \emph{et~al.}}]{E12-06-108A}
\bibinfo{author}{\bibnamefont{Carman}, \bibfnamefont{D.~S.}}, \emph{et~al.},
  \bibinfo{year}{2014}, \bibinfo{title}{\emph{Exclusive {$N^* \to KY$} Studies
  with CLAS12.}}, \bibinfo{note}{{JLab\,12 Experiment: E12-06-108A}}.

\bibitem[{Caso \emph{et~al.}(1998)\citenamefont{Caso}
  \emph{et~al.}}]{Caso:1998}
\bibinfo{author}{\bibnamefont{Caso}, \bibfnamefont{C.}}, \emph{et~al.},
  \bibinfo{year}{1998}, \bibinfo{journal}{Eur. Phys, J. C}
  \textbf{\bibinfo{volume}{3}}, \bibinfo{pages}{1}.

\bibitem[{\citenamefont{Chang} \emph{et~al.}(2009)\citenamefont{Chang,
  Clo{\"e}t, El-Bennich, Kl{\"a}hn, and Roberts}}]{Chang:2009ae}
\bibinfo{author}{\bibnamefont{Chang}, \bibfnamefont{L.}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{El-Bennich}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Kl{\"a}hn}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2009}, \bibinfo{journal}{Chin. Phys. C}
  \textbf{\bibinfo{volume}{33}}, \bibinfo{pages}{1189}.

\bibitem[{\citenamefont{Chang} \emph{et~al.}(2013)\citenamefont{Chang, Roberts,
  and Schmidt}}]{Chang:2012cc}
\bibinfo{author}{\bibnamefont{Chang}, \bibfnamefont{L.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{S.~M.} \bibnamefont{Schmidt}},
  \bibinfo{year}{2013}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{87}}, \bibinfo{pages}{015203}.

\bibitem[{\citenamefont{Chang} \emph{et~al.}(2011)\citenamefont{Chang, Roberts,
  and Tandy}}]{Chang:2011vu}
\bibinfo{author}{\bibnamefont{Chang}, \bibfnamefont{L.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{2011}, \bibinfo{journal}{Chin. J. Phys.}
  \textbf{\bibinfo{volume}{49}}, \bibinfo{pages}{955}.

\bibitem[{\citenamefont{Cheng} \emph{et~al.}(2013)\citenamefont{Cheng,
  Hasenfratz, Petropoulos, and Schaich}}]{Cheng:2013eu}
\bibinfo{author}{\bibnamefont{Cheng}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Hasenfratz}},
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Petropoulos}}, and
  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Schaich}},
  \bibinfo{year}{2013}, \bibinfo{journal}{JHEP} \textbf{\bibinfo{volume}{07}},
  \bibinfo{pages}{061}.

\bibitem[{\citenamefont{Clo{\"e}t and Roberts}(2008)}]{Cloet:2008fw}
\bibinfo{author}{\bibnamefont{Clo{\"e}t}, \bibfnamefont{I.~C.}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2008}, \bibinfo{journal}{PoS}
  \textbf{\bibinfo{volume}{LC2008}}, \bibinfo{pages}{047}.

\bibitem[{\citenamefont{Clo{\"e}t} \emph{et~al.}(2013)\citenamefont{Clo{\"e}t,
  Roberts, and Thomas}}]{Cloet:2013gva}
\bibinfo{author}{\bibnamefont{Clo{\"e}t}, \bibfnamefont{I.~C.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{A.~W.} \bibnamefont{Thomas}},
  \bibinfo{year}{2013}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{101803}.

\bibitem[{\citenamefont{Close and Li}(1990)}]{Close:1989aj}
\bibinfo{author}{\bibnamefont{Close}, \bibfnamefont{F.~E.}}, and
  \bibinfo{author}{\bibfnamefont{Z.-P.} \bibnamefont{Li}},
  \bibinfo{year}{1990}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{42}}, \bibinfo{pages}{2194}.

\bibitem[{\citenamefont{Cornwall}(1982)}]{Cornwall:1981zr}
\bibinfo{author}{\bibnamefont{Cornwall}, \bibfnamefont{J.~M.}},
  \bibinfo{year}{1982}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{1453}.

\bibitem[{\citenamefont{Cornwall and Papavassiliou}(1989)}]{Cornwall:1989gv}
\bibinfo{author}{\bibnamefont{Cornwall}, \bibfnamefont{J.~M.}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Papavassiliou}},
  \bibinfo{year}{1989}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{40}}, \bibinfo{pages}{3474}.

\bibitem[{\citenamefont{Crede and Roberts}(2013)}]{Crede:2013sze}
\bibinfo{author}{\bibnamefont{Crede}, \bibfnamefont{V.}}, and
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Roberts}},
  \bibinfo{year}{2013}, \bibinfo{journal}{Rept. Prog. Phys.}
  \textbf{\bibinfo{volume}{76}}, \bibinfo{pages}{076301}.

\bibitem[{\citenamefont{Cutkosky and Wang}(1990)}]{Cutkosky:1990zh}
\bibinfo{author}{\bibnamefont{Cutkosky}, \bibfnamefont{R.~E.}}, and
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Wang}}, \bibinfo{year}{1990},
  \bibinfo{journal}{Phys. Rev. D} \textbf{\bibinfo{volume}{42}},
  \bibinfo{pages}{235}.

\bibitem[{\citenamefont{Cyrol} \emph{et~al.}(2016)\citenamefont{Cyrol, Fister,
  Mitter, Pawlowski, and Strodthoff}}]{Cyrol:2016tym}
\bibinfo{author}{\bibnamefont{Cyrol}, \bibfnamefont{A.~K.}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Fister}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Mitter}},
  \bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Pawlowski}}, and
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Strodthoff}},
  \bibinfo{year}{2016}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{94}}(\bibinfo{number}{5}), \bibinfo{pages}{054005}.

\bibitem[{\citenamefont{Dalitz and Moorhouse}(1965)}]{DALITZ1965159}
\bibinfo{author}{\bibnamefont{Dalitz}, \bibfnamefont{R.}}, and
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Moorhouse}},
  \bibinfo{year}{1965}, \bibinfo{journal}{Physics Letters}
  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{159}.

\bibitem[{\citenamefont{DeGrand}(2016)}]{DeGrand:2015zxa}
\bibinfo{author}{\bibnamefont{DeGrand}, \bibfnamefont{T.}},
  \bibinfo{year}{2016}, \bibinfo{journal}{Rev. Mod. Phys.}
  \textbf{\bibinfo{volume}{88}}, \bibinfo{pages}{015001}.

\bibitem[{\citenamefont{Deur} \emph{et~al.}(2016)\citenamefont{Deur, Brodsky,
  and de~Teramond}}]{Deur:2016tte}
\bibinfo{author}{\bibnamefont{Deur}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{S.~J.} \bibnamefont{Brodsky}}, and
  \bibinfo{author}{\bibfnamefont{G.~F.} \bibnamefont{de~Teramond}},
  \bibinfo{year}{2016}, \bibinfo{journal}{Prog. Part. Nucl. Phys.}
  \textbf{\bibinfo{volume}{90}}, \bibinfo{pages}{1}.

\bibitem[{\citenamefont{Deur} \emph{et~al.}(2007)\citenamefont{Deur, Burkert,
  Chen, and Korsch}}]{Deur:2005cf}
\bibinfo{author}{\bibnamefont{Deur}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Burkert}},
  \bibinfo{author}{\bibfnamefont{J.-P.} \bibnamefont{Chen}}, and
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Korsch}},
  \bibinfo{year}{2007}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{650}}, \bibinfo{pages}{244}.

\bibitem[{\citenamefont{Deur} \emph{et~al.}(2008)\citenamefont{Deur, Burkert,
  Chen, and Korsch}}]{Deur:2008rf}
\bibinfo{author}{\bibnamefont{Deur}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Burkert}},
  \bibinfo{author}{\bibfnamefont{J.~P.} \bibnamefont{Chen}}, and
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Korsch}},
  \bibinfo{year}{2008}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{665}}, \bibinfo{pages}{349}.

\bibitem[{\citenamefont{Deur} \emph{et~al.}(2014)\citenamefont{Deur, Prok,
  Burkert, Crabb, Girod, Griffioen, Guler, Kuhn, and Kvaltine}}]{Deur:2014vea}
\bibinfo{author}{\bibnamefont{Deur}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{Y.}~\bibnamefont{Prok}},
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Burkert}},
  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Crabb}},
  \bibinfo{author}{\bibfnamefont{F.~X.} \bibnamefont{Girod}},
  \bibinfo{author}{\bibfnamefont{K.~A.} \bibnamefont{Griffioen}},
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Guler}},
  \bibinfo{author}{\bibfnamefont{S.~E.} \bibnamefont{Kuhn}}, and
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Kvaltine}},
  \bibinfo{year}{2014}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{90}}(\bibinfo{number}{1}), \bibinfo{pages}{012009}.

\bibitem[{\citenamefont{Dokshitzer}(1977)}]{Dokshitzer:1977}
\bibinfo{author}{\bibnamefont{Dokshitzer}, \bibfnamefont{Y.~L.}},
  \bibinfo{year}{1977}, \bibinfo{journal}{Sov. Phys. JETP}
  \textbf{\bibinfo{volume}{46}}, \bibinfo{pages}{641}.

\bibitem[{\citenamefont{D{\"o}ring}(2014)}]{Doring:2014qaa}
\bibinfo{author}{\bibnamefont{D{\"o}ring}, \bibfnamefont{M.}},
  \bibinfo{year}{2014}, \bibinfo{journal}{Int. J. Mod. Phys. Conf. Ser.}
  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{1460054}.

\bibitem[{\citenamefont{D{\"o}ring}
  \emph{et~al.}(2009)\citenamefont{D{\"o}ring, Hanhart, Huang, Krewald, and
  Meissner}}]{Doring:2009yv}
\bibinfo{author}{\bibnamefont{D{\"o}ring}, \bibfnamefont{M.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Hanhart}},
  \bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Huang}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Krewald}}, and
  \bibinfo{author}{\bibfnamefont{U.~G.} \bibnamefont{Meissner}},
  \bibinfo{year}{2009}, \bibinfo{journal}{Nucl. Phys. A}
  \textbf{\bibinfo{volume}{829}}, \bibinfo{pages}{170}.

\bibitem[{\citenamefont{Drechsel} \emph{et~al.}(1999)\citenamefont{Drechsel,
  Hanstein, Kamalov, and Tiator}}]{Drechsel:1998hk}
\bibinfo{author}{\bibnamefont{Drechsel}, \bibfnamefont{D.}},
  \bibinfo{author}{\bibfnamefont{O.}~\bibnamefont{Hanstein}},
  \bibinfo{author}{\bibfnamefont{S.~S.} \bibnamefont{Kamalov}}, and
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Tiator}},
  \bibinfo{year}{1999}, \bibinfo{journal}{Nucl. Phys. A}
  \textbf{\bibinfo{volume}{645}}, \bibinfo{pages}{145}.

\bibitem[{\citenamefont{Drechsel} \emph{et~al.}(2007)\citenamefont{Drechsel,
  Kamalov, and Tiator}}]{Drechsel:2007if}
\bibinfo{author}{\bibnamefont{Drechsel}, \bibfnamefont{D.}},
  \bibinfo{author}{\bibfnamefont{S.~S.} \bibnamefont{Kamalov}}, and
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Tiator}},
  \bibinfo{year}{2007}, \bibinfo{journal}{Eur. Phys. J. A}
  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{69}.

\bibitem[{Dugger \emph{et~al.}(2009)\citenamefont{Dugger}
  \emph{et~al.}}]{Dugger:2009pn}
\bibinfo{author}{\bibnamefont{Dugger}, \bibfnamefont{M.}}, \emph{et~al.},
  \bibinfo{year}{2009}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{79}}, \bibinfo{pages}{065206}.

\bibitem[{Durr \emph{et~al.}(2008)\citenamefont{Durr}
  \emph{et~al.}}]{Durr:2008zz}
\bibinfo{author}{\bibnamefont{Durr}, \bibfnamefont{S.}}, \emph{et~al.},
  \bibinfo{year}{2008}, \bibinfo{journal}{Science}
  \textbf{\bibinfo{volume}{322}}, \bibinfo{pages}{1224}.

\bibitem[{\citenamefont{Edwards} \emph{et~al.}(2011)\citenamefont{Edwards,
  Dudek, Richards, and Wallace}}]{Edwards:2011jj}
\bibinfo{author}{\bibnamefont{Edwards}, \bibfnamefont{R.~G.}},
  \bibinfo{author}{\bibfnamefont{J.~J.} \bibnamefont{Dudek}},
  \bibinfo{author}{\bibfnamefont{D.~G.} \bibnamefont{Richards}}, and
  \bibinfo{author}{\bibfnamefont{S.~J.} \bibnamefont{Wallace}},
  \bibinfo{year}{2011}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{84}}, \bibinfo{pages}{074508}.

\bibitem[{\citenamefont{Efimov} \emph{et~al.}(1990)\citenamefont{Efimov,
  Ivanov, and Lyubovitskij}}]{Efimov:1990uz}
\bibinfo{author}{\bibnamefont{Efimov}, \bibfnamefont{G.~V.}},
  \bibinfo{author}{\bibfnamefont{M.~A.} \bibnamefont{Ivanov}}, and
  \bibinfo{author}{\bibfnamefont{V.~E.} \bibnamefont{Lyubovitskij}},
  \bibinfo{year}{1990}, \bibinfo{journal}{Z. Phys. C}
  \textbf{\bibinfo{volume}{47}}, \bibinfo{pages}{583}.

\bibitem[{\citenamefont{Efremov and Radyushkin}(1980)}]{Efremov:1979qk}
\bibinfo{author}{\bibnamefont{Efremov}, \bibfnamefont{A.~V.}}, and
  \bibinfo{author}{\bibfnamefont{A.~V.} \bibnamefont{Radyushkin}},
  \bibinfo{year}{1980}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{94}}, \bibinfo{pages}{245}.

\bibitem[{\citenamefont{Eichmann} \emph{et~al.}(2008)\citenamefont{Eichmann,
  Alkofer, Clo{\"e}t, Krassnigg, and Roberts}}]{Eichmann:2008ae}
\bibinfo{author}{\bibnamefont{Eichmann}, \bibfnamefont{G.}},
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Alkofer}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Krassnigg}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2008}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{042202(R)}.

\bibitem[{\citenamefont{Eichmann} \emph{et~al.}(2010)\citenamefont{Eichmann,
  Alkofer, Krassnigg, and Nicmorus}}]{Eichmann:2009qa}
\bibinfo{author}{\bibnamefont{Eichmann}, \bibfnamefont{G.}},
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Alkofer}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Krassnigg}}, and
  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Nicmorus}},
  \bibinfo{year}{2010}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{104}}, \bibinfo{pages}{201601}.

\bibitem[{\citenamefont{Eichmann} \emph{et~al.}(2009)\citenamefont{Eichmann,
  Clo{\"e}t, Alkofer, Krassnigg, and Roberts}}]{Eichmann:2008ef}
\bibinfo{author}{\bibnamefont{Eichmann}, \bibfnamefont{G.}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Alkofer}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Krassnigg}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2009}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{79}}, \bibinfo{pages}{012202(R)}.

\bibitem[{\citenamefont{Eichmann}
  \emph{et~al.}(2016{\natexlab{a}})\citenamefont{Eichmann, Fischer, and
  Sanchis-Alepuz}}]{Eichmann:2016hgl}
\bibinfo{author}{\bibnamefont{Eichmann}, \bibfnamefont{G.}},
  \bibinfo{author}{\bibfnamefont{C.~S.} \bibnamefont{Fischer}}, and
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Sanchis-Alepuz}},
  \bibinfo{year}{2016}{\natexlab{a}}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{94}}, \bibinfo{pages}{094033}.

\bibitem[{\citenamefont{Eichmann}
  \emph{et~al.}(2016{\natexlab{b}})\citenamefont{Eichmann, Sanchis-Alepuz,
  Williams, Alkofer, and Fischer}}]{Eichmann:2016yit}
\bibinfo{author}{\bibnamefont{Eichmann}, \bibfnamefont{G.}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Sanchis-Alepuz}},
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Williams}},
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Alkofer}}, and
  \bibinfo{author}{\bibfnamefont{C.~S.} \bibnamefont{Fischer}},
  \bibinfo{year}{2016}{\natexlab{b}}, \bibinfo{journal}{Prog. Part. Nucl.
  Phys.} \textbf{\bibinfo{volume}{91}}, \bibinfo{pages}{1}.

\bibitem[{\citenamefont{Engel} \emph{et~al.}(2013)\citenamefont{Engel, Lang,
  Mohler, and Sch{\"a}fer}}]{Engel:2013ig}
\bibinfo{author}{\bibnamefont{Engel}, \bibfnamefont{G.~P.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Lang}},
  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Mohler}}, and
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Sch{\"a}fer}},
  \bibinfo{year}{2013}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{87}}, \bibinfo{pages}{074504}.

\bibitem[{\citenamefont{Gao} \emph{et~al.}(2017)\citenamefont{Gao, Chang, Liu,
  Roberts, and Tandy}}]{Gao:2017mmp}
\bibinfo{author}{\bibnamefont{Gao}, \bibfnamefont{F.}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Chang}},
  \bibinfo{author}{\bibfnamefont{Y.-X.} \bibnamefont{Liu}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{2017}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{96}}, \bibinfo{pages}{034024}.

\bibitem[{\citenamefont{Gattringer and Lang}(2010)}]{Gattringer:2010zz}
\bibinfo{author}{\bibnamefont{Gattringer}, \bibfnamefont{C.}}, and
  \bibinfo{author}{\bibfnamefont{C.~B.} \bibnamefont{Lang}},
  \bibinfo{year}{2010}, \bibinfo{journal}{Lect. Notes Phys.}
  \textbf{\bibinfo{volume}{788}}, \bibinfo{pages}{1}.

\bibitem[{\citenamefont{Gell-Mann}(1964)}]{GellMann:1964nj}
\bibinfo{author}{\bibnamefont{Gell-Mann}, \bibfnamefont{M.}},
  \bibinfo{year}{1964}, \bibinfo{journal}{Phys. Lett.}
  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{214}.

\bibitem[{\citenamefont{Gell-Mann and Low}(1954)}]{GellMann:1954fq}
\bibinfo{author}{\bibnamefont{Gell-Mann}, \bibfnamefont{M.}}, and
  \bibinfo{author}{\bibfnamefont{F.~E.} \bibnamefont{Low}},
  \bibinfo{year}{1954}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{1300}.

\bibitem[{\citenamefont{Gerhardt}(1980)}]{Gerhardt:1980yg}
\bibinfo{author}{\bibnamefont{Gerhardt}, \bibfnamefont{C.}},
  \bibinfo{year}{1980}, \bibinfo{journal}{Z. Phys. C}
  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{311}.

\bibitem[{\citenamefont{Giannini and Santopinto}(2015)}]{Giannini:2015zia}
\bibinfo{author}{\bibnamefont{Giannini}, \bibfnamefont{M.~M.}}, and
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Santopinto}},
  \bibinfo{year}{2015}, \bibinfo{journal}{Chin. J. Phys.}
  \textbf{\bibinfo{volume}{53}}, \bibinfo{pages}{020301}.

\bibitem[{\citenamefont{Glozman and Riska}(1996)}]{Glozman:1995fu}
\bibinfo{author}{\bibnamefont{Glozman}, \bibfnamefont{L.~{\relax Ya}.}}, and
  \bibinfo{author}{\bibfnamefont{D.~O.} \bibnamefont{Riska}},
  \bibinfo{year}{1996}, \bibinfo{journal}{Phys. Rept.}
  \textbf{\bibinfo{volume}{268}}, \bibinfo{pages}{263}.

\bibitem[{\citenamefont{Godfrey and Isgur}(1985)}]{Godfrey:1985xj}
\bibinfo{author}{\bibnamefont{Godfrey}, \bibfnamefont{S.}}, and
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Isgur}},
  \bibinfo{year}{1985}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{32}}, \bibinfo{pages}{189}.

\bibitem[{Gothe \emph{et~al.}(2009)\citenamefont{Gothe}
  \emph{et~al.}}]{E12-09-003}
\bibinfo{author}{\bibnamefont{Gothe}, \bibfnamefont{R.~W.}}, \emph{et~al.},
  \bibinfo{year}{2009}, \bibinfo{title}{\emph{Nucleon Resonance Studies with
  CLAS12.}}, \bibinfo{note}{jLab\,12 Experiment: E12-09-003}.

\bibitem[{\citenamefont{Greenberg}(1964)}]{Greenberg:1964pe}
\bibinfo{author}{\bibnamefont{Greenberg}, \bibfnamefont{O.~W.}},
  \bibinfo{year}{1964}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{598}.

\bibitem[{\citenamefont{Gribov and Lipatov}(1972)}]{Gribov:1972}
\bibinfo{author}{\bibnamefont{Gribov}, \bibfnamefont{V.~N.}}, and
  \bibinfo{author}{\bibfnamefont{L.~N.} \bibnamefont{Lipatov}},
  \bibinfo{year}{1972}, \bibinfo{journal}{Sov. J. Nucl. Phys.}
  \textbf{\bibinfo{volume}{15}}, \bibinfo{pages}{438}.

\bibitem[{\citenamefont{Gromes and Stamatescu}(1976)}]{Gromes:1976cr}
\bibinfo{author}{\bibnamefont{Gromes}, \bibfnamefont{D.}}, and
  \bibinfo{author}{\bibfnamefont{I.~O.} \bibnamefont{Stamatescu}},
  \bibinfo{year}{1976}, \bibinfo{journal}{Nucl. Phys. B}
  \textbf{\bibinfo{volume}{112}}, \bibinfo{pages}{213}.

\bibitem[{\citenamefont{Gross}(2005)}]{Gross:2005kv}
\bibinfo{author}{\bibnamefont{Gross}, \bibfnamefont{D.}}, \bibinfo{year}{2005},
  \bibinfo{journal}{Proc. Nat. Acad. Sci.} \textbf{\bibinfo{volume}{102}},
  \bibinfo{pages}{9099}.

\bibitem[{\citenamefont{Grunberg}(1984)}]{Grunberg:1982fw}
\bibinfo{author}{\bibnamefont{Grunberg}, \bibfnamefont{G.}},
  \bibinfo{year}{1984}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{29}}, \bibinfo{pages}{2315}.

\bibitem[{\citenamefont{Hayakawa} \emph{et~al.}(2011)\citenamefont{Hayakawa,
  Ishikawa, Osaki, Takeda, Uno, and Yamada}}]{Hayakawa:2010yn}
\bibinfo{author}{\bibnamefont{Hayakawa}, \bibfnamefont{M.}},
  \bibinfo{author}{\bibfnamefont{K.~I.} \bibnamefont{Ishikawa}},
  \bibinfo{author}{\bibfnamefont{Y.}~\bibnamefont{Osaki}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Takeda}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Uno}}, and
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Yamada}},
  \bibinfo{year}{2011}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{83}}, \bibinfo{pages}{074509}.

\bibitem[{\citenamefont{Hecht} \emph{et~al.}(2002)\citenamefont{Hecht, Roberts,
  Oettel, Thomas, Schmidt, and Tandy}}]{Hecht:2002ej}
\bibinfo{author}{\bibnamefont{Hecht}, \bibfnamefont{M.~B.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Oettel}},
  \bibinfo{author}{\bibfnamefont{A.~W.} \bibnamefont{Thomas}},
  \bibinfo{author}{\bibfnamefont{S.~M.} \bibnamefont{Schmidt}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{2002}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{65}}, \bibinfo{pages}{055204}.

\bibitem[{\citenamefont{Hey and Kelly}(1983)}]{Hey:1982aj}
\bibinfo{author}{\bibnamefont{Hey}, \bibfnamefont{A.~J.~G.}}, and
  \bibinfo{author}{\bibfnamefont{R.~L.} \bibnamefont{Kelly}},
  \bibinfo{year}{1983}, \bibinfo{journal}{Phys. Rept.}
  \textbf{\bibinfo{volume}{96}}, \bibinfo{pages}{71}.

\bibitem[{\citenamefont{H\mbox{\o{}}gaasen and
  Richard}(1983)}]{Hogaasen:1982rb}
\bibinfo{author}{\bibnamefont{H\mbox{\o{}}gaasen}, \bibfnamefont{H.}}, and
  \bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Richard}},
  \bibinfo{year}{1983}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{124}}, \bibinfo{pages}{520}.

\bibitem[{\citenamefont{H{\"o}ll} \emph{et~al.}(2004)\citenamefont{H{\"o}ll,
  Krassnigg, and Roberts}}]{Holl:2004fr}
\bibinfo{author}{\bibnamefont{H{\"o}ll}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Krassnigg}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2004}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{70}}, \bibinfo{pages}{042203(R)}.

\bibitem[{\citenamefont{Horn and Roberts}(2016)}]{Horn:2016rip}
\bibinfo{author}{\bibnamefont{Horn}, \bibfnamefont{T.}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2016}, \bibinfo{journal}{J. Phys. G.}
  \textbf{\bibinfo{volume}{43}}, \bibinfo{pages}{073001/1}.

\bibitem[{\citenamefont{Isgur}(2000)}]{Isgur:2000ad}
\bibinfo{author}{\bibnamefont{Isgur}, \bibfnamefont{N.}}, \bibinfo{year}{2000},
  in \emph{\bibinfo{booktitle}{{Newport News 2000, Excited nucleons and
  hadronic structure}}}, pp. \bibinfo{pages}{403--422}.

\bibitem[{\citenamefont{Isgur and Karl}(1979)}]{Isgur:1978wd}
\bibinfo{author}{\bibnamefont{Isgur}, \bibfnamefont{N.}}, and
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Karl}}, \bibinfo{year}{1979},
  \bibinfo{journal}{Phys. Rev. D} \textbf{\bibinfo{volume}{19}},
  \bibinfo{pages}{2653}, \bibinfo{note}{[Erratum: Phys. Rev.D23,817(1981)]}.

\bibitem[{\citenamefont{Ishii}(1998)}]{Ishii:1998tw}
\bibinfo{author}{\bibnamefont{Ishii}, \bibfnamefont{N.}}, \bibinfo{year}{1998},
  \bibinfo{journal}{Phys. Lett. B} \textbf{\bibinfo{volume}{431}},
  \bibinfo{pages}{1}.

\bibitem[{Isupov \emph{et~al.}(2017)\citenamefont{Isupov}
  \emph{et~al.}}]{Isupov:2017lnd}
\bibinfo{author}{\bibnamefont{Isupov}, \bibfnamefont{E.~L.}}, \emph{et~al.}
  (\bibinfo{collaboration}{CLAS}), \bibinfo{year}{2017},
  \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{96}}(\bibinfo{number}{2}), \bibinfo{pages}{025209}.

\bibitem[{\citenamefont{Jain and Munczek}(1993)}]{Jain:1993qh}
\bibinfo{author}{\bibnamefont{Jain}, \bibfnamefont{P.}}, and
  \bibinfo{author}{\bibfnamefont{H.~J.} \bibnamefont{Munczek}},
  \bibinfo{year}{1993}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{48}}, \bibinfo{pages}{5403}.

\bibitem[{\citenamefont{Julia-Diaz}
  \emph{et~al.}(2007{\natexlab{a}})\citenamefont{Julia-Diaz, Lee, Matsuyama,
  and Sato}}]{JuliaDiaz:2007kz}
\bibinfo{author}{\bibnamefont{Julia-Diaz}, \bibfnamefont{B.}},
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Matsuyama}}, and
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Sato}},
  \bibinfo{year}{2007}{\natexlab{a}}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{76}}, \bibinfo{pages}{065201}.

\bibitem[{\citenamefont{Julia-Diaz}
  \emph{et~al.}(2007{\natexlab{b}})\citenamefont{Julia-Diaz, Lee, Sato, and
  Smith}}]{JuliaDiaz:2006xt}
\bibinfo{author}{\bibnamefont{Julia-Diaz}, \bibfnamefont{B.}},
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Sato}}, and
  \bibinfo{author}{\bibfnamefont{L.~C.} \bibnamefont{Smith}},
  \bibinfo{year}{2007}{\natexlab{b}}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{75}}, \bibinfo{pages}{015205}.

\bibitem[{\citenamefont{Julia-Diaz and Riska}(2006)}]{JuliaDiaz:2006av}
\bibinfo{author}{\bibnamefont{Julia-Diaz}, \bibfnamefont{B.}}, and
  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Riska}},
  \bibinfo{year}{2006}, \bibinfo{journal}{Nucl.Phys.}
  \textbf{\bibinfo{volume}{A780}}, \bibinfo{pages}{175}.

\bibitem[{\citenamefont{Kamano and Lee}(2012)}]{Kamano:2011ut}
\bibinfo{author}{\bibnamefont{Kamano}, \bibfnamefont{H.}}, and
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{year}{2012}, \bibinfo{journal}{AIP Conf. Proc.}
  \textbf{\bibinfo{volume}{1432}}, \bibinfo{pages}{74}.

\bibitem[{\citenamefont{Kamano} \emph{et~al.}(2010)\citenamefont{Kamano,
  Nakamura, Lee, and Sato}}]{Kamano:2010ud}
\bibinfo{author}{\bibnamefont{Kamano}, \bibfnamefont{H.}},
  \bibinfo{author}{\bibfnamefont{S.~X.} \bibnamefont{Nakamura}},
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}}, and
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Sato}}, \bibinfo{year}{2010},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{81}},
  \bibinfo{pages}{065207}.

\bibitem[{\citenamefont{Kamano} \emph{et~al.}(2013)\citenamefont{Kamano,
  Nakamura, Lee, and Sato}}]{Kamano:2013iva}
\bibinfo{author}{\bibnamefont{Kamano}, \bibfnamefont{H.}},
  \bibinfo{author}{\bibfnamefont{S.~X.} \bibnamefont{Nakamura}},
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}}, and
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Sato}}, \bibinfo{year}{2013},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{88}},
  \bibinfo{pages}{035209}.

\bibitem[{Kim \emph{et~al.}(1998)\citenamefont{Kim}
  \emph{et~al.}}]{Kim:1998kia}
\bibinfo{author}{\bibnamefont{Kim}, \bibfnamefont{J.~H.}}, \emph{et~al.},
  \bibinfo{year}{1998}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{81}}, \bibinfo{pages}{3595}.

\bibitem[{\citenamefont{Koniuk and Isgur}(1980)}]{Koniuk:1979vy}
\bibinfo{author}{\bibnamefont{Koniuk}, \bibfnamefont{R.}}, and
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Isgur}},
  \bibinfo{year}{1980}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{1868},
  \bibinfo{note}{[Erratum: Phys. Rev.D23,818(1981)]}.

\bibitem[{\citenamefont{Lee}(2007)}]{Lee:2007zy}
\bibinfo{author}{\bibnamefont{Lee}, \bibfnamefont{T.~S.~H.}},
  \bibinfo{year}{2007}, \bibinfo{journal}{J. Phys. Conf. Ser.}
  \textbf{\bibinfo{volume}{69}}, \bibinfo{pages}{012013}.

\bibitem[{\citenamefont{Lee}(2013)}]{Lee:2012mea}
\bibinfo{author}{\bibnamefont{Lee}, \bibfnamefont{T.~S.~H.}},
  \bibinfo{year}{2013}, \bibinfo{journal}{AIP Conf. Proc.}
  \textbf{\bibinfo{volume}{1560}}, \bibinfo{pages}{413}.

\bibitem[{\citenamefont{Lepage and Brodsky}(1979)}]{Lepage:1979zb}
\bibinfo{author}{\bibnamefont{Lepage}, \bibfnamefont{G.~P.}}, and
  \bibinfo{author}{\bibfnamefont{S.~J.} \bibnamefont{Brodsky}},
  \bibinfo{year}{1979}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{87}}, \bibinfo{pages}{359}.

\bibitem[{\citenamefont{Lepage and Brodsky}(1980)}]{Lepage:1980fj}
\bibinfo{author}{\bibnamefont{Lepage}, \bibfnamefont{G.~P.}}, and
  \bibinfo{author}{\bibfnamefont{S.~J.} \bibnamefont{Brodsky}},
  \bibinfo{year}{1980}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{2157}.

\bibitem[{\citenamefont{Li} \emph{et~al.}(1992)\citenamefont{Li, Burkert, and
  Li}}]{Li:1991yba}
\bibinfo{author}{\bibnamefont{Li}, \bibfnamefont{Z.-p.}},
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Burkert}}, and
  \bibinfo{author}{\bibfnamefont{Z.-j.} \bibnamefont{Li}},
  \bibinfo{year}{1992}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{46}}, \bibinfo{pages}{70}.

\bibitem[{\citenamefont{Lichtenberg and Tassie}(1967)}]{Lichtenberg:1967zz}
\bibinfo{author}{\bibnamefont{Lichtenberg}, \bibfnamefont{D.~B.}}, and
  \bibinfo{author}{\bibfnamefont{L.~J.} \bibnamefont{Tassie}},
  \bibinfo{year}{1967}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{155}}, \bibinfo{pages}{1601}.

\bibitem[{\citenamefont{Lichtenberg}
  \emph{et~al.}(1968)\citenamefont{Lichtenberg, Tassie, and
  Keleman}}]{Lichtenberg:1968zz}
\bibinfo{author}{\bibnamefont{Lichtenberg}, \bibfnamefont{D.~B.}},
  \bibinfo{author}{\bibfnamefont{L.~J.} \bibnamefont{Tassie}}, and
  \bibinfo{author}{\bibfnamefont{P.~J.} \bibnamefont{Keleman}},
  \bibinfo{year}{1968}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{167}}, \bibinfo{pages}{1535}.

\bibitem[{\citenamefont{Lin and Cohen}(2012)}]{Lin:2011da}
\bibinfo{author}{\bibnamefont{Lin}, \bibfnamefont{H.-W.}}, and
  \bibinfo{author}{\bibfnamefont{S.~D.} \bibnamefont{Cohen}},
  \bibinfo{year}{2012}, \bibinfo{journal}{AIP Conf. Proc.}
  \textbf{\bibinfo{volume}{1432}}, \bibinfo{pages}{305}.

\bibitem[{\citenamefont{Lin} \emph{et~al.}(2008)\citenamefont{Lin, Cohen,
  Edwards, and Richards}}]{Lin:2008qv}
\bibinfo{author}{\bibnamefont{Lin}, \bibfnamefont{H.-W.}},
  \bibinfo{author}{\bibfnamefont{S.~D.} \bibnamefont{Cohen}},
  \bibinfo{author}{\bibfnamefont{R.~G.} \bibnamefont{Edwards}}, and
  \bibinfo{author}{\bibfnamefont{D.~G.} \bibnamefont{Richards}},
  \bibinfo{year}{2008}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{78}}, \bibinfo{pages}{114508}.

\bibitem[{\citenamefont{Liu}(2017)}]{Liu:2016rwa}
\bibinfo{author}{\bibnamefont{Liu}, \bibfnamefont{K.-F.}},
  \bibinfo{year}{2017}, \bibinfo{journal}{Int. J. Mod. Phys. E}
  \textbf{\bibinfo{volume}{26}}, \bibinfo{pages}{1740016}.

\bibitem[{\citenamefont{Liu} \emph{et~al.}(2014)\citenamefont{Liu, Chen, Gong,
  Sufian, Sun, and Li}}]{Liu:2014jua}
\bibinfo{author}{\bibnamefont{Liu}, \bibfnamefont{K.-F.}},
  \bibinfo{author}{\bibfnamefont{Y.}~\bibnamefont{Chen}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Gong}},
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Sufian}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Sun}}, and
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Li}}, \bibinfo{year}{2014},
  \bibinfo{journal}{PoS} \textbf{\bibinfo{volume}{LATTICE2013}},
  \bibinfo{pages}{507}.

\bibitem[{\citenamefont{L{\"o}ring}
  \emph{et~al.}(2001)\citenamefont{L{\"o}ring, Metsch, and
  Petry}}]{Loring:2001kx}
\bibinfo{author}{\bibnamefont{L{\"o}ring}, \bibfnamefont{U.}},
  \bibinfo{author}{\bibfnamefont{B.~C.} \bibnamefont{Metsch}}, and
  \bibinfo{author}{\bibfnamefont{H.~R.} \bibnamefont{Petry}},
  \bibinfo{year}{2001}, \bibinfo{journal}{Eur. Phys. J. A}
  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{395}.

\bibitem[{\citenamefont{Lu} \emph{et~al.}(2017)\citenamefont{Lu, Chen, Roberts,
  Segovia, Xu, and Zong}}]{Lu:2017cln}
\bibinfo{author}{\bibnamefont{Lu}, \bibfnamefont{Y.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Chen}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Segovia}},
  \bibinfo{author}{\bibfnamefont{S.-S.} \bibnamefont{Xu}}, and
  \bibinfo{author}{\bibfnamefont{H.-S.} \bibnamefont{Zong}},
  \bibinfo{year}{2017}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{96}}, \bibinfo{pages}{015208}.

\bibitem[{\citenamefont{Mahbub} \emph{et~al.}(2012)\citenamefont{Mahbub,
  Kamleh, Leinweber, Moran, and Williams}}]{Mahbub:2010rm}
\bibinfo{author}{\bibnamefont{Mahbub}, \bibfnamefont{M.~S.}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Kamleh}},
  \bibinfo{author}{\bibfnamefont{D.~B.} \bibnamefont{Leinweber}},
  \bibinfo{author}{\bibfnamefont{P.~J.} \bibnamefont{Moran}}, and
  \bibinfo{author}{\bibfnamefont{A.~G.} \bibnamefont{Williams}},
  \bibinfo{year}{2012}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{707}}, \bibinfo{pages}{389}.

\bibitem[{\citenamefont{Marciano and Pagels}(1978)}]{Marciano:1977su}
\bibinfo{author}{\bibnamefont{Marciano}, \bibfnamefont{W.~J.}}, and
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Pagels}},
  \bibinfo{year}{1978}, \bibinfo{journal}{Phys. Rept.}
  \textbf{\bibinfo{volume}{36}}, \bibinfo{pages}{137}.

\bibitem[{\citenamefont{Marciano and Pagels}(1979)}]{Marciano:1979wa}
\bibinfo{author}{\bibnamefont{Marciano}, \bibfnamefont{W.~J.}}, and
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Pagels}},
  \bibinfo{year}{1979}, \bibinfo{journal}{Nature}
  \textbf{\bibinfo{volume}{279}}, \bibinfo{pages}{479}.

\bibitem[{\citenamefont{Maris}(2002)}]{Maris:2002yu}
\bibinfo{author}{\bibnamefont{Maris}, \bibfnamefont{P.}}, \bibinfo{year}{2002},
  \bibinfo{journal}{Few Body Syst.} \textbf{\bibinfo{volume}{32}},
  \bibinfo{pages}{41}.

\bibitem[{\citenamefont{Maris}(2004)}]{Maris:2004bp}
\bibinfo{author}{\bibnamefont{Maris}, \bibfnamefont{P.}}, \bibinfo{year}{2004},
  \bibinfo{journal}{Few Body Syst.} \textbf{\bibinfo{volume}{35}},
  \bibinfo{pages}{117}.

\bibitem[{\citenamefont{Maris and Roberts}(1997)}]{Maris:1997tm}
\bibinfo{author}{\bibnamefont{Maris}, \bibfnamefont{P.}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{1997}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{56}}, \bibinfo{pages}{3369}.

\bibitem[{\citenamefont{Maris} \emph{et~al.}(1998)\citenamefont{Maris, Roberts,
  and Tandy}}]{Maris:1997hd}
\bibinfo{author}{\bibnamefont{Maris}, \bibfnamefont{P.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{1998}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{420}}, \bibinfo{pages}{267}.

\bibitem[{\citenamefont{Maris and Tandy}(1999)}]{Maris:1999nt}
\bibinfo{author}{\bibnamefont{Maris}, \bibfnamefont{P.}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{1999}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{60}}, \bibinfo{pages}{055214}.

\bibitem[{\citenamefont{Matsuyama} \emph{et~al.}(2007)\citenamefont{Matsuyama,
  Sato, and Lee}}]{Matsuyama:2006rp}
\bibinfo{author}{\bibnamefont{Matsuyama}, \bibfnamefont{A.}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Sato}}, and
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{year}{2007}, \bibinfo{journal}{Phys. Rept.}
  \textbf{\bibinfo{volume}{439}}, \bibinfo{pages}{193}.

\bibitem[{\citenamefont{Miller}(2007)}]{Miller:2007uy}
\bibinfo{author}{\bibnamefont{Miller}, \bibfnamefont{G.~A.}},
  \bibinfo{year}{2007}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{99}}, \bibinfo{pages}{112001}.

\bibitem[{\citenamefont{Mokeev} \emph{et~al.}(2016)\citenamefont{Mokeev,
  Aznauryan, Burkert, and Gothe}}]{Mokeev:2015moa}
\bibinfo{author}{\bibnamefont{Mokeev}, \bibfnamefont{V.~I.}},
  \bibinfo{author}{\bibfnamefont{I.}~\bibnamefont{Aznauryan}},
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Burkert}}, and
  \bibinfo{author}{\bibfnamefont{R.}~\bibnamefont{Gothe}},
  \bibinfo{year}{2016}, \bibinfo{journal}{EPJ Web Conf.}
  \textbf{\bibinfo{volume}{113}}, \bibinfo{pages}{01013}.

\bibitem[{Mokeev \emph{et~al.}(2012)\citenamefont{Mokeev}
  \emph{et~al.}}]{Mokeev:2012vsa}
\bibinfo{author}{\bibnamefont{Mokeev}, \bibfnamefont{V.~I.}}, \emph{et~al.},
  \bibinfo{year}{2012}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{035203}.

\bibitem[{Mokeev \emph{et~al.}(2016)\citenamefont{Mokeev}
  \emph{et~al.}}]{Mokeev:2015lda}
\bibinfo{author}{\bibnamefont{Mokeev}, \bibfnamefont{V.~I.}}, \emph{et~al.},
  \bibinfo{year}{2016}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{93}}, \bibinfo{pages}{025206}.

\bibitem[{\citenamefont{Moorhouse}(1966)}]{PhysRevLett.16.772}
\bibinfo{author}{\bibnamefont{Moorhouse}, \bibfnamefont{R.~G.}},
  \bibinfo{year}{1966}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{16}}, \bibinfo{pages}{772}.

\bibitem[{\citenamefont{Nambu}(2011)}]{Nambu:2011zz}
\bibinfo{author}{\bibnamefont{Nambu}, \bibfnamefont{Y.}}, \bibinfo{year}{2011},
  \bibinfo{journal}{AIP Conf. Proc.} \textbf{\bibinfo{volume}{1388}},
  \bibinfo{pages}{86}.

\bibitem[{\citenamefont{Oettel} \emph{et~al.}(2000)\citenamefont{Oettel,
  Pichowsky, and von Smekal}}]{Oettel:1999gc}
\bibinfo{author}{\bibnamefont{Oettel}, \bibfnamefont{M.}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Pichowsky}}, and
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{von Smekal}},
  \bibinfo{year}{2000}, \bibinfo{journal}{Eur. Phys. J. A}
  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{251}.

\bibitem[{\citenamefont{Pascalutsa}
  \emph{et~al.}(2007)\citenamefont{Pascalutsa, Vanderhaeghen, and
  Yang}}]{Pascalutsa:2006up}
\bibinfo{author}{\bibnamefont{Pascalutsa}, \bibfnamefont{V.}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Vanderhaeghen}}, and
  \bibinfo{author}{\bibfnamefont{S.~N.} \bibnamefont{Yang}},
  \bibinfo{year}{2007}, \bibinfo{journal}{Phys. Rept.}
  \textbf{\bibinfo{volume}{437}}, \bibinfo{pages}{125}.

\bibitem[{Patrignani \emph{et~al.}(2016)\citenamefont{Patrignani}
  \emph{et~al.}}]{Olive:2016xmw}
\bibinfo{author}{\bibnamefont{Patrignani}, \bibfnamefont{C.}}, \emph{et~al.},
  \bibinfo{year}{2016}, \bibinfo{journal}{Chin. Phys. C}
  \textbf{\bibinfo{volume}{40}}(\bibinfo{number}{10}), \bibinfo{pages}{100001}.

\bibitem[{\citenamefont{Pilaftsis}(1997)}]{Pilaftsis:1996fh}
\bibinfo{author}{\bibnamefont{Pilaftsis}, \bibfnamefont{A.}},
  \bibinfo{year}{1997}, \bibinfo{journal}{Nucl. Phys. B}
  \textbf{\bibinfo{volume}{487}}, \bibinfo{pages}{467}.

\bibitem[{\citenamefont{Politzer}(2005)}]{Politzer:2005kc}
\bibinfo{author}{\bibnamefont{Politzer}, \bibfnamefont{H.~D.}},
  \bibinfo{year}{2005}, \bibinfo{journal}{Proc. Nat. Acad. Sci.}
  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{7789}.

\bibitem[{\citenamefont{Qin} \emph{et~al.}(2011)\citenamefont{Qin, Chang, Liu,
  Roberts, and Wilson}}]{Qin:2011dd}
\bibinfo{author}{\bibnamefont{Qin}, \bibfnamefont{S.-X.}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Chang}},
  \bibinfo{author}{\bibfnamefont{Y.-X.} \bibnamefont{Liu}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{D.~J.} \bibnamefont{Wilson}},
  \bibinfo{year}{2011}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{84}}, \bibinfo{pages}{042202(R)}.

\bibitem[{\citenamefont{Qin} \emph{et~al.}(2012)\citenamefont{Qin, Chang, Liu,
  Roberts, and Wilson}}]{Qin:2011xq}
\bibinfo{author}{\bibnamefont{Qin}, \bibfnamefont{S.-X.}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Chang}},
  \bibinfo{author}{\bibfnamefont{Y.-X.} \bibnamefont{Liu}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{D.~J.} \bibnamefont{Wilson}},
  \bibinfo{year}{2012}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{85}}, \bibinfo{pages}{035202}.

\bibitem[{\citenamefont{Raya} \emph{et~al.}(2016)\citenamefont{Raya, Chang,
  Bashir, Cobos-Martinez, Guti{\'e}rrez-Guerrero, Roberts, and
  Tandy}}]{Raya:2015gva}
\bibinfo{author}{\bibnamefont{Raya}, \bibfnamefont{K.}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Chang}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Bashir}},
  \bibinfo{author}{\bibfnamefont{J.~J.} \bibnamefont{Cobos-Martinez}},
  \bibinfo{author}{\bibfnamefont{L.~X.} \bibnamefont{Guti{\'e}rrez-Guerrero}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{P.~C.} \bibnamefont{Tandy}},
  \bibinfo{year}{2016}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{93}}, \bibinfo{pages}{074017}.

\bibitem[{\citenamefont{Reinhardt}(1990)}]{Reinhardt:1989rw}
\bibinfo{author}{\bibnamefont{Reinhardt}, \bibfnamefont{H.}},
  \bibinfo{year}{1990}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{244}}, \bibinfo{pages}{316}.

\bibitem[{\citenamefont{Richard}(1992)}]{Richard:1992uk}
\bibinfo{author}{\bibnamefont{Richard}, \bibfnamefont{J.~M.}},
  \bibinfo{year}{1992}, \bibinfo{journal}{Phys. Rept.}
  \textbf{\bibinfo{volume}{212}}, \bibinfo{pages}{1}.

\bibitem[{Ripani \emph{et~al.}(2003)\citenamefont{Ripani}
  \emph{et~al.}}]{Ripani:2002ss}
\bibinfo{author}{\bibnamefont{Ripani}, \bibfnamefont{M.}}, \emph{et~al.},
  \bibinfo{year}{2003}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{91}}, \bibinfo{pages}{022002}.

\bibitem[{\citenamefont{Roberts}(2015)}]{Roberts:2015dea}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{C.~D.}},
  \bibinfo{year}{2015}, \bibinfo{journal}{J. Phys. Conf. Ser.}
  \textbf{\bibinfo{volume}{630}}(\bibinfo{number}{1}), \bibinfo{pages}{012051}.

\bibitem[{\citenamefont{Roberts}(2016)}]{Roberts:2015lja}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{C.~D.}},
  \bibinfo{year}{2016}, \bibinfo{journal}{J. Phys. Conf. Ser.}
  \textbf{\bibinfo{volume}{706}}, \bibinfo{pages}{022003}.

\bibitem[{\citenamefont{Roberts}(2017)}]{Roberts:2016vyn}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{C.~D.}},
  \bibinfo{year}{2017}, \bibinfo{journal}{Few Body Syst.}
  \textbf{\bibinfo{volume}{58}}, \bibinfo{pages}{5}.

\bibitem[{\citenamefont{Roberts}
  \emph{et~al.}(2013{\natexlab{a}})\citenamefont{Roberts, Holt, and
  Schmidt}}]{Roberts:2013mja}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{C.~D.}},
  \bibinfo{author}{\bibfnamefont{R.~J.} \bibnamefont{Holt}}, and
  \bibinfo{author}{\bibfnamefont{S.~M.} \bibnamefont{Schmidt}},
  \bibinfo{year}{2013}{\natexlab{a}}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{727}}, \bibinfo{pages}{249}.

\bibitem[{\citenamefont{Roberts and Segovia}(2016)}]{Roberts:2016dnb}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{C.~D.}}, and
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Segovia}},
  \bibinfo{year}{2016}, \bibinfo{journal}{Few Body Syst.}
  \textbf{\bibinfo{volume}{57}}(\bibinfo{number}{11}), \bibinfo{pages}{1067}.

\bibitem[{\citenamefont{Roberts and Williams}(1994)}]{Roberts:1994dr}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{C.~D.}}, and
  \bibinfo{author}{\bibfnamefont{A.~G.} \bibnamefont{Williams}},
  \bibinfo{year}{1994}, \bibinfo{journal}{Prog. Part. Nucl. Phys.}
  \textbf{\bibinfo{volume}{33}}, \bibinfo{pages}{477}.

\bibitem[{\citenamefont{Roberts}
  \emph{et~al.}(2013{\natexlab{b}})\citenamefont{Roberts, Kamleh, and
  Leinweber}}]{Roberts:2013ipa}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{D.~S.}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Kamleh}}, and
  \bibinfo{author}{\bibfnamefont{D.~B.} \bibnamefont{Leinweber}},
  \bibinfo{year}{2013}{\natexlab{b}}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{725}}, \bibinfo{pages}{164}.

\bibitem[{\citenamefont{Roberts} \emph{et~al.}(2011)\citenamefont{Roberts,
  Bashir, Guti{\'e}rrez-Guerrero, Roberts, and Wilson}}]{Roberts:2011wy}
\bibinfo{author}{\bibnamefont{Roberts}, \bibfnamefont{H.~L.~L.}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Bashir}},
  \bibinfo{author}{\bibfnamefont{L.~X.} \bibnamefont{Guti{\'e}rrez-Guerrero}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{D.~J.} \bibnamefont{Wilson}},
  \bibinfo{year}{2011}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{83}}, \bibinfo{pages}{065206}.

\bibitem[{\citenamefont{Rojas} \emph{et~al.}(2014)\citenamefont{Rojas,
  El-Bennich, and de~Melo}}]{Rojas:2014aka}
\bibinfo{author}{\bibnamefont{Rojas}, \bibfnamefont{E.}},
  \bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{El-Bennich}}, and
  \bibinfo{author}{\bibfnamefont{J.~P. B.~C.} \bibnamefont{de~Melo}},
  \bibinfo{year}{2014}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{90}}, \bibinfo{pages}{074025}.

\bibitem[{\citenamefont{R{\"o}nchen}
  \emph{et~al.}(2013)\citenamefont{R{\"o}nchen, D{\"o}ring, Huang, Haberzettl,
  Haidenbauer, Hanhart, Krewald, Meissner, and Nakayama}}]{Ronchen:2012eg}
\bibinfo{author}{\bibnamefont{R{\"o}nchen}, \bibfnamefont{D.}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{D{\"o}ring}},
  \bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Huang}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Haberzettl}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Haidenbauer}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Hanhart}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Krewald}},
  \bibinfo{author}{\bibfnamefont{U.~G.} \bibnamefont{Meissner}}, and
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Nakayama}},
  \bibinfo{year}{2013}, \bibinfo{journal}{Eur. Phys. J. A}
  \textbf{\bibinfo{volume}{49}}, \bibinfo{pages}{44}.

\bibitem[{\citenamefont{Roper}(1964)}]{Roper:1964zza}
\bibinfo{author}{\bibnamefont{Roper}, \bibfnamefont{L.~D.}},
  \bibinfo{year}{1964}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{340}.

\bibitem[{\citenamefont{Roper} \emph{et~al.}(1965)\citenamefont{Roper, Wright,
  and Feld}}]{PhysRev.138.B190}
\bibinfo{author}{\bibnamefont{Roper}, \bibfnamefont{L.~D.}},
  \bibinfo{author}{\bibfnamefont{R.~M.} \bibnamefont{Wright}}, and
  \bibinfo{author}{\bibfnamefont{B.~T.} \bibnamefont{Feld}},
  \bibinfo{year}{1965}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{138}}, \bibinfo{pages}{B190}.

\bibitem[{\citenamefont{Sanchis-Alepuz}
  \emph{et~al.}(2014)\citenamefont{Sanchis-Alepuz, Fischer, and
  Kubrak}}]{Sanchis-Alepuz:2014wea}
\bibinfo{author}{\bibnamefont{Sanchis-Alepuz}, \bibfnamefont{H.}},
  \bibinfo{author}{\bibfnamefont{C.~S.} \bibnamefont{Fischer}}, and
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Kubrak}},
  \bibinfo{year}{2014}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{733}}, \bibinfo{pages}{151}.

\bibitem[{\citenamefont{Sannino}(2009)}]{Sannino:2009za}
\bibinfo{author}{\bibnamefont{Sannino}, \bibfnamefont{F.}},
  \bibinfo{year}{2009}, \bibinfo{journal}{Acta Phys. Polon. B}
  \textbf{\bibinfo{volume}{40}}, \bibinfo{pages}{3533}.

\bibitem[{\citenamefont{Sato and Lee}(1996)}]{Sato:1996gk}
\bibinfo{author}{\bibnamefont{Sato}, \bibfnamefont{T.}}, and
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{year}{1996}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{54}}, \bibinfo{pages}{2660}.

\bibitem[{\citenamefont{Sato and Lee}(2001)}]{Sato:2000jf}
\bibinfo{author}{\bibnamefont{Sato}, \bibfnamefont{T.}}, and
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{year}{2001}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{63}}, \bibinfo{pages}{055201}.

\bibitem[{\citenamefont{Segovia}
  \emph{et~al.}(2014{\natexlab{a}})\citenamefont{Segovia, Chen, Clo{\"e}t,
  Roberts, Schmidt, and Wan}}]{Segovia:2013uga}
\bibinfo{author}{\bibnamefont{Segovia}, \bibfnamefont{J.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Chen}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{author}{\bibfnamefont{S.~M.} \bibnamefont{Schmidt}}, and
  \bibinfo{author}{\bibfnamefont{S.-L.} \bibnamefont{Wan}},
  \bibinfo{year}{2014}{\natexlab{a}}, \bibinfo{journal}{Few Body Syst.}
  \textbf{\bibinfo{volume}{55}}, \bibinfo{pages}{1}.

\bibitem[{\citenamefont{Segovia} \emph{et~al.}(2013)\citenamefont{Segovia,
  Chen, Roberts, and Wan}}]{Segovia:2013rca}
\bibinfo{author}{\bibnamefont{Segovia}, \bibfnamefont{J.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Chen}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{S.-L.} \bibnamefont{Wan}},
  \bibinfo{year}{2013}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{88}}, \bibinfo{pages}{032201(R)}.

\bibitem[{\citenamefont{Segovia}
  \emph{et~al.}(2014{\natexlab{b}})\citenamefont{Segovia, Clo{\"e}t, Roberts,
  and Schmidt}}]{Segovia:2014aza}
\bibinfo{author}{\bibnamefont{Segovia}, \bibfnamefont{J.}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{S.~M.} \bibnamefont{Schmidt}},
  \bibinfo{year}{2014}{\natexlab{b}}, \bibinfo{journal}{Few Body Syst.}
  \textbf{\bibinfo{volume}{55}}, \bibinfo{pages}{1185}.

\bibitem[{\citenamefont{Segovia}
  \emph{et~al.}(2015{\natexlab{a}})\citenamefont{Segovia, El-Bennich, Rojas,
  Clo{\"e}t, Roberts, Xu, and Zong}}]{Segovia:2015hra}
\bibinfo{author}{\bibnamefont{Segovia}, \bibfnamefont{J.}},
  \bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{El-Bennich}},
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Rojas}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{author}{\bibfnamefont{S.-S.} \bibnamefont{Xu}}, and
  \bibinfo{author}{\bibfnamefont{H.-S.} \bibnamefont{Zong}},
  \bibinfo{year}{2015}{\natexlab{a}}, \bibinfo{journal}{Phys. Rev. Lett.}
  \textbf{\bibinfo{volume}{115}}(\bibinfo{number}{17}),
  \bibinfo{pages}{171801}.

\bibitem[{\citenamefont{Segovia and Roberts}(2016)}]{Segovia:2016zyc}
\bibinfo{author}{\bibnamefont{Segovia}, \bibfnamefont{J.}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2016}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{94}}, \bibinfo{pages}{042201(R)}.

\bibitem[{\citenamefont{Segovia}
  \emph{et~al.}(2015{\natexlab{b}})\citenamefont{Segovia, Roberts, and
  Schmidt}}]{Segovia:2015ufa}
\bibinfo{author}{\bibnamefont{Segovia}, \bibfnamefont{J.}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}}, and
  \bibinfo{author}{\bibfnamefont{S.~M.} \bibnamefont{Schmidt}},
  \bibinfo{year}{2015}{\natexlab{b}}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{750}}, \bibinfo{pages}{100}.

\bibitem[{\citenamefont{{Shen, C.-P.}}(2016)}]{ChengpingFB21}
\bibinfo{author}{\bibnamefont{{Shen, C.-P.}}}, \bibinfo{year}{2016},
  \bibinfo{journal}{EPJ Web Conf.} \textbf{\bibinfo{volume}{113}},
  \bibinfo{pages}{01016}.

\bibitem[{\citenamefont{Suzuki} \emph{et~al.}(2010)\citenamefont{Suzuki,
  Julia-Diaz, Kamano, Lee, Matsuyama, and Sato}}]{Suzuki:2009nj}
\bibinfo{author}{\bibnamefont{Suzuki}, \bibfnamefont{N.}},
  \bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{Julia-Diaz}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Kamano}},
  \bibinfo{author}{\bibfnamefont{T.~S.~H.} \bibnamefont{Lee}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Matsuyama}}, and
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Sato}}, \bibinfo{year}{2010},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{104}},
  \bibinfo{pages}{042302}.

\bibitem[{\citenamefont{Tiator} \emph{et~al.}(2004)\citenamefont{Tiator,
  Drechsel, Kamalov, Giannini, Santopinto, and Vassallo}}]{Tiator:2003uu}
\bibinfo{author}{\bibnamefont{Tiator}, \bibfnamefont{L.}},
  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Drechsel}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Kamalov}},
  \bibinfo{author}{\bibfnamefont{M.~M.} \bibnamefont{Giannini}},
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Santopinto}}, and
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Vassallo}},
  \bibinfo{year}{2004}, \bibinfo{journal}{Eur. Phys. J. A}
  \textbf{\bibinfo{volume}{19}}, \bibinfo{pages}{55},
  \bibinfo{note}{[,55(2003)]}.

\bibitem[{\citenamefont{Tiator and Vanderhaeghen}(2009)}]{Tiator:2008kd}
\bibinfo{author}{\bibnamefont{Tiator}, \bibfnamefont{L.}}, and
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Vanderhaeghen}},
  \bibinfo{year}{2009}, \bibinfo{journal}{Phys. Lett. B}
  \textbf{\bibinfo{volume}{672}}, \bibinfo{pages}{344}.

\bibitem[{Villano \emph{et~al.}(2009)\citenamefont{Villano}
  \emph{et~al.}}]{Villano:2009sn}
\bibinfo{author}{\bibnamefont{Villano}, \bibfnamefont{A.~N.}}, \emph{et~al.},
  \bibinfo{year}{2009}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{80}}, \bibinfo{pages}{035203}.

\bibitem[{\citenamefont{Warns} \emph{et~al.}(1990)\citenamefont{Warns,
  Schroder, Pfeil, and Rollnik}}]{Warns:1989ie}
\bibinfo{author}{\bibnamefont{Warns}, \bibfnamefont{M.}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Schroder}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Pfeil}}, and
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Rollnik}},
  \bibinfo{year}{1990}, \bibinfo{journal}{Z. Phys. C}
  \textbf{\bibinfo{volume}{45}}, \bibinfo{pages}{627}.

\bibitem[{\citenamefont{Wilczek}(2000)}]{Wilczek:2000ih}
\bibinfo{author}{\bibnamefont{Wilczek}, \bibfnamefont{F.}},
  \bibinfo{year}{2000}, \bibinfo{journal}{Phys. Today}
  \textbf{\bibinfo{volume}{53N8}}, \bibinfo{pages}{22}.

\bibitem[{\citenamefont{Wilczek}(2005)}]{Wilczek:2005az}
\bibinfo{author}{\bibnamefont{Wilczek}, \bibfnamefont{F.}},
  \bibinfo{year}{2005}, \bibinfo{journal}{Proc. Nat. Acad. Sci.}
  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{8403}.

\bibitem[{\citenamefont{Wilczek}(2012)}]{Wilczek:2012sb}
\bibinfo{author}{\bibnamefont{Wilczek}, \bibfnamefont{F.}},
  \bibinfo{year}{2012}, \bibinfo{journal}{Central Eur. J. Phys.}
  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{1021}.

\bibitem[{\citenamefont{Wilson} \emph{et~al.}(2012)\citenamefont{Wilson,
  Clo{\"e}t, Chang, and Roberts}}]{Wilson:2011aa}
\bibinfo{author}{\bibnamefont{Wilson}, \bibfnamefont{D.~J.}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{L.}~\bibnamefont{Chang}}, and
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{year}{2012}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{85}}, \bibinfo{pages}{025205}.

\bibitem[{\citenamefont{Wilson}(1974)}]{Wilson:1974sk}
\bibinfo{author}{\bibnamefont{Wilson}, \bibfnamefont{K.~G.}},
  \bibinfo{year}{1974}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{2445}.

\bibitem[{\citenamefont{Xu} \emph{et~al.}(2015)\citenamefont{Xu, Chen,
  Clo{\"e}t, Roberts, Segovia, and Zong}}]{Xu:2015kta}
\bibinfo{author}{\bibnamefont{Xu}, \bibfnamefont{S.-S.}},
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Chen}},
  \bibinfo{author}{\bibfnamefont{I.~C.} \bibnamefont{Clo{\"e}t}},
  \bibinfo{author}{\bibfnamefont{C.~D.} \bibnamefont{Roberts}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Segovia}}, and
  \bibinfo{author}{\bibfnamefont{H.-S.} \bibnamefont{Zong}},
  \bibinfo{year}{2015}, \bibinfo{journal}{Phys. Rev. D}
  \textbf{\bibinfo{volume}{92}}, \bibinfo{pages}{114034}.

\bibitem[{\citenamefont{Yang and Ping}(arXiv:2017.10nnn [hep-ph])}]{Yang:2017}
\bibinfo{author}{\bibnamefont{Yang}, \bibfnamefont{G.}}, and
  \bibinfo{author}{\bibfnamefont{J.-L.} \bibnamefont{Ping}},
  \bibinfo{year}{arXiv:2017.10nnn [hep-ph]}, \bibinfo{note}{{\emph{The order of
  1P and 2S states of baryons in the chiral quark model}}}.

\bibitem[{\citenamefont{Zweig}(1964)}]{Zweig:1981pd}
\bibinfo{author}{\bibnamefont{Zweig}, \bibfnamefont{G.}}, \bibinfo{year}{1964},
  \bibinfo{note}{{\emph{An $SU(3)$ model for strong interaction symmetry and
  its breaking. Parts 1 and 2} (CERN Reports No.\ 8182/TH.\ 401 and No.\
  8419/TH.\ 412)}}.

\end{thebibliography}


\end{document}
