\begin{thebibliography}{12}

\bibitem{Kelly86}
E. J. Kelly, ``An adaptive detection algorithm," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 22, no. 1, pp. 115-127, Mar. 1986.

%\bibitem{Kelly89}
% E. J. Kelly and K. Forsythe, ``Adaptive detection and parameter estimation for multidimensional signal models," Lincoln %Lab., Mass. Inst., Technol., Lexington, USA, Tech. Rep. no. 848, Apr. 1989.

\bibitem{Scharf94}
L. L. Scharf and B. Friedlander, ``Matched subspace detectors," \emph{IEEE Trans. Signal Process.,} vol. 42, no. 8, pp. 2146-2157, Aug. 1994.

\bibitem{Kraut01}
S. Kraut, L. L. Scharf, and L. T. McWhorterer, ``Adaptive subspace detectors," \emph{IEEE Trans. Signal Process.,} vol. 49, no. 1, pp. 1-16, Jan. 2001.

\bibitem{Besson_Dec06}
O. Besson, L. L. Scharf, and S. Kraut, ``Adaptive detection of a signal known only to lie
on a line in a known subspace, when primary and secondary data are partially homogeneous,"
 \emph{IEEE Trans. Signal Process.,} vol. 54, no. 12, pp. 4698-4705, Dec. 2006.

 \bibitem{Besson_Jul06}
 O. Besson, ``Detection of a signal in linear subspace with bounded mismatch," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 42,
no. 2, pp. 1131-1139, July 2006.

\bibitem{Bandiera_June07}
F. Bandiera, O. Besson, D. Orlando, G. Ricci, and L. L. Scharf, ``GLRT-based direction detectors in homogeneous noise and subspace interference," \emph{IEEE Trans. Signal Process.,} vol. 55, no. 6, pp. 2386-2394, June 2007.

\bibitem{Bandiera_Apr07}
F. Bandiera, A. D. Maio, A. S. Greco, and G. Ricci, ``Adaptive radar detection of distributed targets in homogeneous and partially homogeneous
noise plus subspace interference," \emph{IEEE Trans. Signal Process.,} vol. 55, no. 4, pp. 1223-1237, Apr. 2007.

\bibitem{Svensson_Dec11}
A. Svensson and A. Jakobsson, ``Adaptive
detection of a partly known signal corrupted by strong interference," \emph{IEEE Signal Process. Lett.,} vol. 18, no. 12, pp. 729-732, Dec. 2011.

\bibitem{Aubry_Jul14}
A. Aubry, A. D. Maio, D. Orlando, and M. Piezzo, ``Adaptive
detection of
point-like targets in the presence of homogeneous clutter and subspace interference," \emph{IEEE Signal Process. Lett.,} vol. 21, no. 7,
pp. 848-852, July 2014.

\bibitem{Ramprashad96}
S. Ramprashad, T. W. Parks, and R. Shenoy, ``Signal modeling and detection using cone classes," \emph{IEEE Trans. Signal Process.,} vol. 4, no. 2,
pp. 329-338, Feb. 1996.

\bibitem{DeMaio05}
 A. D. Maio, ``Robust adaptive radar detection in the presence of steering vector mismatches," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 41,
no. 4, pp. 1322-1337, Oct. 2005.

\bibitem{Bandiera_June07_conic}
 F. Bandiera, A. D. Maio, and G. Ricci, ``Adaptive CFAR radar detection with conic rejection," \emph{IEEE Trans. Signal Process.,} vol. 55,
no. 6, pp. 2533-2541, June 2007.

 \bibitem{Bandiera_Sep09}
 F. Bandiera, D. Orlando, and G. Ricci, ``CFAR detection strategies for distributed targets under conic constraints," \emph{IEEE Trans. Signal Process.,} vol. 57,
no. 9, pp. 3305-3316, Sep. 2009.

\bibitem{DeMaio09}
 A. D. Maio, S. D. Nicola, Y. Huang, S. Zhang, and A. Farina, ``Adaptive detection and estimation in the presence of useful signal and interference mismatches," \emph{IEEE Trans. Signal Process.,} vol. 57,
no. 2, pp. 436-450, Feb. 2009.

\bibitem{Conte87}
E. Conte and M. Longo, ``Characterisation of radar clutter as a spherically invariant random process,"
\emph{IEE Proc.,} vol. 134, no. 2, pp. 191-197, Apr. 1987.

\bibitem{Ward90}
K. D. Ward, C. J. Baker, and S. Watt, ``Maritime surveillance radar-Part I: Radar scattering from the ocean surface," \emph{IEE Proc.,} vol. 137, no. 2, pp. 51-62, Apr. 1990.

\bibitem{Conte95}
E. Conte, M. Lops, and G. Ricci, ``Asymptotically optimum radar detection in compound-Gaussian clutter," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 31, no. 2, pp. 617-625, Apr. 1995.

\bibitem{Conte96}
E. Conte, M. Lops, and G. Ricci, ``Adaptive matched filter detection in spherically invariant noise," \emph{IEEE Signal Process. Lett.,} vol. 3, no. 8, pp. 248-250, Aug. 1996.

\bibitem{Conte98}
E. Conte, M. Lops, and G. Ricci, ``Adaptive detection schemes in compound Gaussian clutter," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 34, no. 4, pp. 1058-1069, Oct. 1998.

\bibitem{Gini99}
F. Gini and M. V. Greco, ``Supoptimum approach to adaptive coherent radar detection in compound-Gaussian clutter," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 35, no. 3, pp. 1095-1104, July 1999.

\bibitem{Gerlach_Jul99}
K. Gerlach, ``Spatially distributed target detection in non-Gaussian clutter," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 35, no. 3, pp. 926-934, July 1999.


\bibitem{Conte_Apr02}
E. Conte, A. D. Maio, and G. Ricci, ``CFAR detection of distributed targets in non-Gaussian disturbance," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 38, no. 2, pp. 612-621, Apr. 2002.

\bibitem{Conte_Apr02_cov_est}
E. Conte, A. D. Maio, and G. Ricci, ``Covariance matrix estimation for adaptive CFAR detection in compound Gaussian clutter," \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 38, no. 2, pp. 415-426, Apr. 2002.

\bibitem{Conte_Aug02}
E. Conte, A. D. Maio, and G. Ricci, ``Recursive estimation of the covariance matrix of a compound Gaussian process and its
application to adaptive CFAR detection," \emph{IEEE Trans. Signal Process.,} vol. 50, no. 8, pp. 1908-1915, Aug. 2002.

\bibitem{Pascal08}
F. Pascal, Y. Chitour, J. P. Ovarlez, and P. Forster, ``Covariance structure maximum-likelihood estimates
in compound Gaussian noise: Existence and algorithms analysis," \emph{IEEE Trans. Signal Process.,} vol. 56, no. 1, pp. 34-48, Jan. 2008.

\bibitem{He10}
Y. He, T. Jian, F. Su, C. Qu, and X. Gu, ``Novel range-spread target detectors in non-Gaussian clutter,"
\emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 46, no. 3, pp. 1312-1327, July 2010.

\bibitem{Jian12}
 Y. He, T. Jian, F. Su, C. Qu, and D. Ping, ``Adaptive detection application of covariance matrix estimator for correlated non-Gaussian clutter,"
 \emph{IEEE Trans. Signal Process.,} vol. 46, no. 4, pp. 2108-2117, Oct. 2010.

 \bibitem{He10}
T. Jian, Y. He, F. Su, C. Qu, and D. Ping, ``Cascaded detector for range-spread target in non-Gaussian clutter,"
 \emph{IEEE Trans. Signal Process.,} vol. 48, no. 2, pp. 1713-1725, Apr. 2012.

\bibitem{Rabaste_Jul14}
O. Rabaste and N. Trouve, ``Geometrical design of radar detectors in moderately impulsive noise,"
 \emph{IEEE Trans. Signal Process.,} vol. 50, no. 3, pp. 1938-1954, July 2014.

\bibitem{DeMaio11}
 A. D. Maio, Y. Huang, D. P. Palomar, S. Zhang, and A. Farina, ``Fractional QCQP with applications in ML steering direction estimation for radar detection," \emph{IEEE Trans. Signal Process.,} vol. 59,
no. 1, pp. 172-185, Jan. 2011.

\bibitem{Roh06}
T. Roh and L. Vandenberghe, ``Discrete transforms, semidefinite programming, and sum-of-squares representations of nonnegative polynomials," \emph{SIAM J. Optim.,} vol. 16, no. 4, pp. 939-964, Jan. 2006.

\bibitem{Watt90}
S. Watt, C. J. Baker, and K. D. Ward, ``Maritime surveillance radar-Part 2: Detection performance prediction in sea clutter," \emph{IEE Proc.,} vol. 137, no. 2, pp. 63-72, Apr. 1990.

\bibitem{Ward81}
K. D. Ward, ``Compound representation of high resolution sea clutter," \emph{Electron. Lett.,} vol. 17, no. 16, pp. 561-563, Aug. 1981.

\bibitem{Conte01}
 E. Conte, A. D. Maio, and G. Ricci, ``GLRT-based adaptive detection algorithms for range spread targets," \emph{IEEE Trans. Signal Process.,} vol. 49,
no. 7, pp. 1336-1348, July 2001.


\bibitem{Luo10}
Z. Q. Luo, A. M. C. So, Y. Ye, and S. Zhang, ``Semidefinite relaxation of quadratic optimization problems,"
\emph{IEEE Signal Process. Mag.,} vol. 27, no. 3, pp. 20-34, May 2010.

\bibitem{DeMaio14}
A. D. Maio and Y. Huang, ``New results on fractional QCQP with applications to radar steering direction estimation," \emph{IEEE Signal Process. Lett.,} vol. 21, no. 7, pp. 895-898, July 2014.

\bibitem{Davidson02}
 T. N. Davidson, Z. Q. Luo, and J. F. Sturm, ``Linear matrix inequality formulation of spectral mask constraints with applications
 to FIR filter design,"
 \emph{IEEE Trans. Signal Process.,} vol. 50,
no. 11, pp. 2702-2715, Nov. 2002.

\bibitem{DeMaio10}
 A. D. Maio, S. D. Nicola, A. Farina, and S. Iommelli, ``Adaptive detection of a signal with angle uncertainty," \emph{IET Radar, Sonar, Navigation,} vol. 4,
no. 4, pp. 537-547, Aug. 2010.

\bibitem{Rangaswamy95}
M. Rangaswamy, D. Weiner, and A. Ozturk, ``Computer generation of correlated non-Gaussian radar clutter,"
 \emph{IEEE Trans. Aerosp. Electron. Syst.,} vol. 31,
no. 1, pp. 106-116, Jan. 1995.





\end{thebibliography} 